### 2021-06-08 thesis/index.Rmd
Merge branch 'main' of github.com:rfl-urbaniak/BayesBias

---
title: "Conceptual and methodological problems with bias detection and avoidance in natural language processing"
author: "Alicja Dobrzeniecka"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_book: 
    keep_tex: yes
    toc_depth: 6
    latex_engine: xelatex
    extra_dependencies: ["todonotes"]
  bookdown::gitbook:
    lib_dir: "book_assets"
includes:
    in_header:
      - Rafal_latex6.sty
documentclass: book
classoptions: dvipsnames, usenames
site: bookdown::bookdown_site
bibliography: [../references/bibliographyRafal.bib]
biblio-style: 
link-citations: yes
fontsize: 12pt
linestretch: 1.5
geometry: left=3.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm, asymmetric, includeheadfoot
csl: apa-5th-edition.csl
---

<!-- github-repo: rstudio/bookdown-demo -->

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(knitr)

options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
```

<!-- # Preface {-} -->

<!-- Test  -->

<!-- testsds -->

<!-- \begin{align}  -->
<!--   f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k} -->
<!--   (\#eq:binom) -->
<!-- \end{align}  -->

<!-- This is a citation [@diamond90] which uses keys from the bib file listed in the preamble. -->

<!-- Equation \@ref(eq:binom)^[This is a footnote containing a double citation [@diamond90; @dahlmanNakedStat2020; @Zhao2019Gender].] -->

<!-- Note that chapter files are found and compiled automatically, but the file names have to contain chapter numbers first. For instance, we used `01-intro.Rmd`, placed in the same folder. Observe how we included r code inline. -->

<!-- ```{r fig-margin, fig.margin=TRUE} -->
<!-- plot(cars) -->
<!-- ``` -->