### 2021-06-10 paperDraft/paperDraft.tex

\documentclass[10pt,dvipsnames,enabledeprecatedfontcommands]{scrartcl}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={A bayesian method of cosine-based word2vec bias},
            pdfauthor={Alicja Dobrzeniecka and Rafal Urbaniak},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

%\documentclass{article}

% %packages
 \usepackage{booktabs}

\usepackage{multirow}

\usepackage{graphicx}
\usepackage{longtable}
\usepackage{ragged2e}
\usepackage{etex}
%\usepackage{yfonts}
\usepackage{marvosym}
\usepackage[notextcomp]{kpfonts}
\usepackage{nicefrac}
\newcommand*{\QED}{\hfill \footnotesize {\sc Q.e.d.}}
\usepackage{floatrow}

\usepackage[textsize=footnotesize]{todonotes}
%\linespread{1.5}

\setlength{\parindent}{10pt}
\setlength{\parskip}{1pt}

%language
\usepackage{times}
\usepackage{t1enc}
%\usepackage[utf8x]{inputenc}
%\usepackage[polish]{babel}
%\usepackage{polski}

\usepackage{mathptmx}
\usepackage[scaled=0.88]{helvet}

%AMS
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}

\usepackage{geometry}
 \geometry{a4paper,left=35mm,top=20mm,}

%environments
\newtheorem{fact}{Fact}

%abbreviations
\newcommand{\ra}{\rangle}
\newcommand{\la}{\langle}
\newcommand{\n}{\neg}
\newcommand{\et}{\wedge}
\newcommand{\jt}{\rightarrow}
\newcommand{\ko}[1]{\forall  #1\,}
\newcommand{\ro}{\leftrightarrow}
\newcommand{\exi}[1]{\exists\, {_{#1}}}
\newcommand{\pr}[1]{\mathsf{P}(#1)}
\newcommand{\cost}{\mathsf{cost}}

\newcommand{\odds}{\mathsf{Odds}}
\newcommand{\ind}{\mathsf{Ind}}
\newcommand{\nf}[2]{\nicefrac{#1\,}{#2}}
\newcommand{\R}[1]{\texttt{#1}}
\newcommand{\prr}[1]{\mbox{$\mathtt{P}_{prior}(#1)$}}
\newcommand{\prp}[1]{\mbox{$\mathtt{P}_{posterior}(#1)$}}

\newtheorem{q}{\color{blue}Question}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

%technical intermezzo
%---------------------

\newcommand{\intermezzoa}{
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Starts}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}\nopagebreak 
	}

\newcommand{\intermezzob}{\nopagebreak 
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Ends}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}
	}
%--------------------

\newtheorem*{reply*}{Reply}
\usepackage{enumitem}
\newcommand{\question}[1]{\begin{enumerate}[resume,leftmargin=0cm,labelsep=0cm,align=left]
\item #1
\end{enumerate}}

\usepackage{float}

% \setbeamertemplate{blocks}[rounded][shadow=true]
% \setbeamertemplate{itemize items}[ball]
% \AtBeginPart{}
% \AtBeginSection{}
% \AtBeginSubsection{}
% \AtBeginSubsubsection{}
% \setlength{\emergencystretch}{0em}
% \setlength{\parskip}{0pt}

\usepackage[authoryear]{natbib}

%\bibliographystyle{apalike}

\title{A bayesian method of cosine-based word2vec bias}
\author{Alicja Dobrzeniecka and Rafal Urbaniak}
\date{}

\begin{document}
\maketitle

test

\section{Cosine-based measures of
bias}\label{cosine-based-measures-of-bias}

One of the first measures in the discussion has been developed by
Bolukbasi, Chang, Zou, Saligrama, \& Kalai (2016). There, the gender
bias of a word \(w\) is understood as its projection on the gender
direction \(\vec{w} \cdot (\overrightarrow{he} - \overrightarrow{she})\)
(the gender direction is the top principal compontent of ten gender pair
difference vectors). The underlying idea is that no bias is present if
non-explicitly gendered words are in equal distance to both elements in
all explicitly gendered pairs. Given the (ideally) gender netural words
\(N\) and the gender direction \(g\) the direct direct gender bias is
defined as the average distance of the words in \(N\) from \(g\) (\(c\)
is a parameter determining how strict we want to be):

\begin{align}
\mathsf{directBias_c(N,g)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},g)\vert^c}{\vert N \vert }
\end{align}

The use of projections has been ciriticized for instance by Gonen \&
Goldberg (2019), who point out that while gender-direction might be an
indicator of bias, it is only one possible manifestation of it, and
reducing a projection of words might be insufficient. For instance,
``math'' and ``delicate'' might be in equal distance to both explicitly
gendered words while being closer to quite different stereotypical
attribute words. Further, the authors point out that most word pairs
preserve similarity under debiasing meant to minimize projection-based
bias.\footnote{Bolukbasi et al. (2016) use also another method which
  involves analogies and their evaluations by human users on Mechanical
  Turk. It is discussed and criticized in (Nissim, Noord, \& Goot,
  2020).}

To measure bias in word embeddings, Caliskan, Bryson, \& Narayanan
(2017) proposed the Word Embedding Association Test (WEAT). The idea is
that the measure of biases between two sets of target words, \(X\) and
\(Y\), (we call them protected words) should be quantified in terms of
the cosine similarity between the protected words and attribute words
coming from two sets of stereotype attribute words, \(A\) and \(B\)
(we'll call them attributes). For instance, \(X\) might be a set of male
names, \(Y\) a set of female names, \(A\) might contain stereotypically
male-related career words, and \(B\) stereotypically female-related
family words. WEAT is a modification of the Implicit Association Test
(IAT) (Nosek, Banaji, \& Greenwald, 2002) used in psychology and uses
almost the same word sets, allowing for a \emph{prima facie} sensible
comparison with bias in humans. Let \(f\) be a distance or similarity
measure (usually, cosine similarity). The association difference for a
term \(t\) is:

\begin{align}
s(t,A,B) & = \frac{\sum_{a\in A}f(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}f(t,b)}{\vert B\vert}
\end{align}

\noindent then, the association difference between \(A\) a \(B\) is:

\begin{align}
s(X,Y,A,B) & = \sum_{x\in X} s(x,A,B) -  \sum_{y\in Y} s(y,A,B)
\end{align}

\noindent
\(s(X,Y,A,B)\) is the statistic used in the signifcance test, and the
\(p\)-value obtained by bootstrapping: it is the frequency of
\(s(X_i,Y_i,A,B)>s(X,Y,A,B)\) for all equally sized partitions
\(X_i, Y_i\) of \(X\cup Y\). The effect size is computed by normalizing
the difference in means as follows:

\begin{align}
bias(A,B) & = \frac{
\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
}
\end{align}

Caliskan et al. (2017) show that significant biases---thus measured---
similar to the ones discovered by IAT can be discovered in word
embeddings. Lauscher \& Glavas (2019) extended the methodology to a
multilingual and cross-lingual setting, arguing that using Euclidean
distance instead of similarty does not make much difference, while the
bias effects vary greatly accross embedding models (interestingly, with
social media-text trained embeddings being less biased than those based
on Wikipedia).

A similar methodology is employed by Garg, Schiebinger, Jurafsky, \& Zou
(2018), who employ word embeddings trained od corpora from different
decades to study the shifts in various biases. For instance, to compute
the occupational embeddings bias for women the authors first compute the
average vector of vector emeddings of words that represent women (e.g.
``she'', ``female''), then calculate the Euclidean distance between this
mean vector and words for occupations. Then they take the mean of these
distances and subtract from it the analogously obtained mean for the
average vector of vector embeddings of words that represent men.
Formally they take the relative norm distance between \(X\) and \(Y\) to
be:

\begin{align}
\textsf{relative norm distance} & = \sum_{v_m\in M} \vert \vert v_m - v_X\vert \vert_2 - \vert v_m - v_Y\vert \vert_2
\end{align}

\noindent where the norm used is Euclidean, and \(v_X\) and \(x_Y\) are
average vectors for sets \(X\) and \(Y\) respectively.

There is a problem with taking means of means in the calculations,
though. By pre-averaging data we throw away information about sample
size. For instance, think about proportions: 10 out of 20 and 2 out of 4
give the same mean, but you would obtain more information by making the
former observation rather than by making the latter. Connectedly, when
we pre-average, we remove variation, and so pre-averaging tend to
manufacture false confidence.

To illustrate let's employ the formulas used by Caliskan et al. (2017)
and go through an example. Conceptually, all such tests come up with
rather short lists of protected words and rather short lists of
stereotypical attributes. Clearly, these are not complete list. So let's
treat them as samples from richer pools of stereotypical predicates and
let's take the uncertainty and variation involved seriously.

Consider a situation in which there are two protected classes,
\(X=\{t_1,t_2\}\) and \(Y=\{t_3,t_4\}\) and two five-element sets \(A\)
and \(B\).

First, we play around with a scenario in which all the protected terms
are on average equally distant from both sets (\(\mu =1\)) with standard
deviation of \(.05\). First, let's randomly draw distances and plot the
results with group means plotted as vertical lines.

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\NormalTok{t1 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t2 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t3 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t4 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-2-1} \end{center}

When you look at the datapoints, do you have the impression of a strong
bias here? We wouldn't think so. But now let's run the calculations from
(Caliskan et al., 2017).

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{s <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{ (table)\{ }\KeywordTok{mean}\NormalTok{(table}\OperatorTok{$}\NormalTok{A) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(table}\OperatorTok{$}\NormalTok{B)\}}
\NormalTok{factor <-}\StringTok{ }\KeywordTok{sd}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\KeywordTok{s}\NormalTok{(t1),}\KeywordTok{s}\NormalTok{(t2),}\KeywordTok{s}\NormalTok{(t3),}\KeywordTok{s}\NormalTok{(t4)))}
\NormalTok{numerator <-}\StringTok{  }\KeywordTok{mean}\NormalTok{(}\KeywordTok{s}\NormalTok{(t1),}\KeywordTok{s}\NormalTok{(t2)) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{s}\NormalTok{(t3),}\KeywordTok{s}\NormalTok{(t4))}
\NormalTok{bias <-}\StringTok{ }\NormalTok{numerator }\OperatorTok{/}\StringTok{ }\NormalTok{factor}
\NormalTok{bias}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.825005
\end{verbatim}

\normalsize

This should make us pause. We know these were points randomly drawn from
distributions where there is no difference in means. Yet, the calculated
effect size is 1.82, whereas the largest effect size reported by
Caliskan et al. (2017) is 1.81!

Now, let's try sampling from distributions where there in fact is a
difference in means. Terms from \(X\) are on average .8 away from \(A\)
(and \(1\) from \(B\)), while terms from \(Y\) are \(.8\) away to \(B\)
(and 1 to \(A\)). The standard deviation is 0.1 in all the cases. There
is a clear difference between \(X\) and \(Y\) and quick visual
inspection should tell us so.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\NormalTok{t1 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{8}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t2 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{8}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t3 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{8}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t4 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{A  =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\DataTypeTok{B =} \KeywordTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{8}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\normalsize

\vspace{1mm} \footnotesize

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-5-1} \end{center}

\normalsize

Is this clear difference mirrored in the calculations?

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{factor <-}\StringTok{ }\KeywordTok{sd}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\KeywordTok{s}\NormalTok{(t1),}\KeywordTok{s}\NormalTok{(t2),}\KeywordTok{s}\NormalTok{(t3),}\KeywordTok{s}\NormalTok{(t4)))}
\NormalTok{numerator <-}\StringTok{  }\KeywordTok{mean}\NormalTok{(}\KeywordTok{s}\NormalTok{(t1),}\KeywordTok{s}\NormalTok{(t2)) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{s}\NormalTok{(t3),}\KeywordTok{s}\NormalTok{(t4))}
\NormalTok{bias <-}\StringTok{ }\NormalTok{numerator }\OperatorTok{/}\StringTok{ }\NormalTok{factor}
\NormalTok{bias}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -1.444026
\end{verbatim}

\normalsize

The absolute value of the effect size is smaller than in the previous
case!

\newpage

\noindent \huge  \textbf{Appendix} \normalsize

\noindent \Large \textbf{Word lists, including human and neutral predicates}
\normalsize

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\vspace{-3mm}

\hypertarget{refs}{}
\hypertarget{ref-bolukbasi2016man}{}
Bolukbasi, T., Chang, K.-W., Zou, J., Saligrama, V., \& Kalai, A.
(2016). Man is to computer programmer as woman is to homemaker?
Debiasing word embeddings.

\hypertarget{ref-Caliskan2017semanticsBiases}{}
Caliskan, A., Bryson, J. J., \& Narayanan, A. (2017). Semantics derived
automatically from language corpora contain human-like biases.
\emph{Science}, \emph{356}(6334), 183--186. American Association for the
Advancement of Science (AAAS). Retrieved from
\url{https://doi.org/10.1126/science.aal4230}

\hypertarget{ref-Garg2018years}{}
Garg, N., Schiebinger, L., Jurafsky, D., \& Zou, J. (2018). Word
embeddings quantify 100 years of gender and ethnic stereotypes.
\emph{Proceedings of the National Academy of Sciences}, \emph{115}(16),
E3635--E3644. Proceedings of the National Academy of Sciences. Retrieved
from \url{https://doi.org/10.1073/pnas.1720347115}

\hypertarget{ref-Gonen2019lipstick}{}
Gonen, H., \& Goldberg, Y. (2019). Lipstick on a pig: Debiasing methods
cover up systematic gender biases in word embeddings but do not remove
them. In \emph{Proceedings of the 2019 conference of the north American
chapter of the association for computational linguistics: Human language
technologies, volume 1 (long and short papers)} (pp. 609--614).
Minneapolis, Minnesota: Association for Computational Linguistics.
Retrieved from \url{https://www.aclweb.org/anthology/N19-1061}

\hypertarget{ref-Lauscher2019multidimensional}{}
Lauscher, A., \& Glavas, G. (2019). Are we consistently biased?
Multidimensional analysis of biases in distributional word vectors.
\emph{CoRR}, \emph{abs/1904.11783}. Retrieved from
\url{http://arxiv.org/abs/1904.11783}

\hypertarget{ref-Nissim2020fair}{}
Nissim, M., Noord, R. van, \& Goot, R. van der. (2020). Fair is better
than sensational: Man is to doctor as woman is to doctor.
\emph{Computational Linguistics}, \emph{46}(2), 487--497. MIT Press -
Journals. Retrieved from \url{https://doi.org/10.1162/coli_a_00379}

\hypertarget{ref-Nosek2002harvesting}{}
Nosek, B. A., Banaji, M. R., \& Greenwald, A. G. (2002). Harvesting
implicit group attitudes and beliefs from a demonstration web site.
\emph{Group Dynamics: Theory, Research, and Practice}, \emph{6}(1),
101--115. American Psychological Association (APA). Retrieved from
\url{https://doi.org/10.1037/1089-2699.6.1.101}

\end{document}