### 2021-06-20 thesis/04-discussion.Rmd
Adds final discussion

---
output:
  pdf_document: default
  html_document: default
---
# Discussion and summary

<!-- Bayesian method -->

We propose the use of Bayesian methods to measure uncertainty in bias detection. There are a few advantages of this method. One of them is that including uncertainty enables one to directly observe the influence of sample sizes. Analyzing individual words and connection coefficients, one may notice how `neutral` words have smaller uncertainty intervals and `different` or `associated` quite the opposite. One of the reasons for such outcome is that we used approximately 230 neutral words and only between 11-25 (the number varies from class to class) stereotypical attributes from @Manzini2019blackToCriminal article. In our approach, we also pay attention to the distribution and details regarding anomalous values. With the use of simple visualizations that we introduced before, we were able to indicate suspicious cosine distance values. Additionally, we compare in details how the cosine distance values and uncertainty change after the debiasing. One can verify then how the individual vectors changed and if it is what was expected. Our analysis with the use of Bayesian method gave us new ideas and hypothesis concerning not only the bias detection method but also the efficiency of debiasing itself.

<!-- General insights - estimated coefficients mean -->

We created a summary table for each of the datasets: Reddit, Reddit Debiased, and Google word embeddings. Let us first analyse the general observations from estimated coefficients mean introduced in 3.6. DATASET-LEVEL COEFFICIENTS. For Google embeddings the HPDI for all classes coefficients (associated, different, human, and none) has an interval that includes zero. This can lead one to a conclusion that the impact for associated, different, human and neutral attribute is, when averaged, quite similar. This indicates how including the uncertainty may change the use and interpretation of @Manzini2019blackToCriminal MAC metric. It seems that if one focuses only on differences between means of means, it is too simplistic. In case of Reddit word embeddings the situation is similar although HPDI interval is below 0 for Gender class when looking at `associated` and `different` mean coefficient. This can suggest that there is indeed slightly stronger impact of these attributes. One should also notice how in general `associated` and `different` coefficients have quite similar HPDI range, the highest observed absolute difference is equal to only 0.1. This suggests that again the impact of associated attributes and difference ones is not clear at first sight when looking at averaged coefficients. Finally, let's compare the HPDI intervals for Reddit and Reddit debiased datasets. For Religion and Race dataset there is a minor shift (in absolute values the highest change is equal to approximately 0.1) of the mean coefficients towards zero. However for Gender dataset there is no significant change. This is of course the general look at the data, let's now analyze the individual words.

<!-- Details from the table -->

In Reddit table one can observe that for Religion class the cosine distance results for the associated attributes are for approximately 60% of the words close to human attributes as well. This can suggest that in some cases words concerning humans can have higher similarity with some protected words independently if they are stereotypical or neutral-human words. One should also pay attention to the fact that for all of the `associated` and `different` attributes in Religion class the uncertainty interval overlaps at some point. What is even more surprising is that for protected words, such as "torah"  the `associated` attribute has the the cosine distance slightly over 1, which means no positive similarity! If the protected words that we chose do not have high similarity with harmful associated attributes, then one should consider at least three scenarios. The first one is that the choice of the protected words and attributes may be corrupted. The second one is that the metric is not able to catch the hidden bias properly. The third one is that there is actually no bias between the words. Regardless of which scenario one considers, it is essential to take a look at the individual values before averaging them or aggregating in other ways. It seems that using Bayesian method can enhance the process of verifying the hypotheses concerning the choice of protected words and attributes.

Surprisingly in Gender class one can observe high cosine similarity values between some female stereotypical professions and male protected words. If a word stereotypically associated with females has low cosine distance to male protected words, then one should investigate the issue further. The reasons for this unexpected cosine distance may lie in the frequency of appearance of the protected words in the raw data. Some of the groups (like Muslim people or females) may have less representation in the data that is taken as an input for word embeddings. Therefore, if we assume that the MAC detects actually co-occurrence only, it makes sense that they can have high similarity with associated attributes and lower similarity with different ones. At the same time male protected words and other religions can have high similarity with different attributes because they have greater representation in the dataset in general and occur close to much more concepts. Cosine distance seems to capture the information on the co-occurrence of words and not on the semantic similarity strictly speaking.

In Google table one may notice interesting observations as well. Although GoogleNews dataset is larger, trained on different data sources and with the use of more dimensions, some of the results are similar to the ones obtained for the Reddit corpus. For Religion case almost all of the values for `associated` class have an intersection with `different` class as well. In the case of Race it is 100% of the available words. This indicates again that it is not clear how the metric should be used. Quite different situation takes place for Gender data where the similarity between `associated` and `different` is present mostly for the male protected words. This means that male words have high similarity with both male stereotypical professions and female ones. However, for females the similarity is high mostly only for the female stereotypical professions. This observation could not be make when using MAC metric only as it requires investigating the individual words.

When analyzing the Reddit debiased results one should ask a question of the change of the cosine distance values. It seems that not all of the protected words are treated equally when performing debiasing. In religion data the values for `associated` class moved towards 0 for most of the words except the ones for Islam religion where still the `associated` class has quite lower cosine distance than the `different` one. Similar situation takes place in Gender data where female words have still much lower cosine distance values for `associated` class than for the `different` one. In the case of males it is mostly almost the same interval for `associated` and `different`. As the Gender data is quite specific as the attributes are not harmful adjectives but (in an ideal world) neutral professions, the aim (if we follow MAC assumptions) should actually be to make the cosine distance same for both female and male protected words. However, as we pointed out the situation after debiasing can sometimes be better only for one protected group which is not the desired outcome. In the case of Religion data even after debiasing, all of the protected words have intersections between `associated` and `different` class and similarity greater than 0. As all of the attributes in religion data are negative and harmful stereotypes, one should not aim at making the distance between protected words and `associated`, and `different` class the same but rather to move it towards 0.

Let's summarize the results of the bias detection methods analysis. As it was presented above, using the bias detection methods without involving the uncertainty may bring the risk of limited insight into the data. One should be concerned with the limitations of the metric that evaluates the debiasing by the mean of the averaged mean approach. Additionally, one cannot be sure if the bias is still preserved after debiasing. The fact that all of the cosine distances for protected words and harmful attributes moved slightly to the right, does not mean that the bias is removed. It is shown in articles such as @Gonen2019Lipstick, that the bias can hide in the vector geometry and preserve even after applying popular debiasing methods. One of the reasons for that may be the fact that relative differences between words may be preserved even after debiasing. Another thing is that the choice of word lists used to verify the metric effectiveness is not well justified. The sample size is very small which leads to large HPDI. Some of the protected words and attributes originate from old articles and it seems that the bias phenomenon is quite dynamic and the data should be as up-to-date as it is possible. One could for example research the IAT examples to verify the effectiveness of MAC metric. One should remember that there is no clear interpretation for the values obtained with MAC metric from @Manzini2019blackToCriminal. One may assume that if the cosine distance is close to 1 then it is a desired outcome as it means, according to cosine distance assumptions, that there is almost no similarity between the words. However what does it mean to be close to 0? If the averaged cosine distance is equal to 0.8, then should we still debias it? It is unclear what the criteria are. On one hand, it seems to be beneficial when the outcome is simplified as it is easier to compare results with one value per set. On the other hand, it is prone to misunderstanding of how to interpret the results and what threshold to assume.

The bottom line is that if we want to take bias seriously, so should we approach the uncertainty involved in our estimations. There is no replacement for proper statistical evaluation that does not discard information about the uncertainty involved, larger word lists are needed, and visualization of the results for particular protected classes provides much better guidance than chasing a single metric based on a means of means.

Bibliography:

<!--

Articles:
  1. "Are We Consistently Biased? Multidimensional Analysis of Biases in Distributional Word Vectors"
  https://www.aclweb.org/anthology/S19-1010.pdf 
  2. Summary what causes bias short article
  [source https://kawine.github.io/blog/nlp/2019/09/23/bias.html]
  3. Men Also Like Shopping: Reducing Gender Bias Amplification using Corpus-level Constraint
  4. Bias in Bios: A Case Study of Semantic Representation Bias in a High-Stakes Setting
  -->