### 2021-06-22 thesis/_book/_main.tex
not much

\documentclass[12pt,]{book}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1.5}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Conceptual and methodological problems with bias detection and avoidance in natural language processing},
            pdfauthor={Alicja Dobrzeniecka},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[left=3.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm, asymmetric,
includeheadfoot]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{todonotes}

\title{Conceptual and methodological problems with bias detection and avoidance
in natural language processing}
\author{Alicja Dobrzeniecka}
\date{2021-06-21}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{5}
\tableofcontents
}
\chapter{Introduction}\label{introduction}

Natural language processing (NLP) is a subfield of computer science that
processes and analyzes language in text and speech with the use of
modern programming methods. It has practical applications in everyday
life as it concerns tasks such as email filters, smart assistants,
search results, language translations, text analytics and so on. Models
used to accomplish these tasks need a lot of data to learn from. This
data originates from human activities and historical recordings such as
texts, messages or speeches. It turns out that in the learning process
these models can learn implicit biases that reflect harmful
stereotypical thinking still present in modern societies. There are many
different types of models in NLP depending on a task that they are
supposed to solve. However, all of them need as an input words
represented by means of numbers and this is accomplished with word
embeddings. The word embeddings are usually assigned the values based on
the context in which the words appear. This means that the input data
can have an enormous influence on the outcome. The biases seem to have
their primary source in the way the words are assigned numerical values.
One can find methods that aim at identifying and measuring hidden biases
and/or try to remove them by for example modifying the word embeddings
directly.

One can find various definitions trying to capture what bias and
fairness actually are. With the choice of the definition, implications
into the real-life applications may change as well. Mehrabi, Morstatter,
Saxena, Lerman, \& Galstyan
(\protect\hyperlink{ref-Mehrabi2019Survey}{2019}) mark out that there
exist different types of biases such as historical bias, representation
bias or measurement bias (the list is long). This indicates how complex
the issue of bias is. Without the proper understanding and awareness of
the problem, people are prone to unconsciously sustain the bias. Mehrabi
et al. (\protect\hyperlink{ref-Mehrabi2019Survey}{2019}) also
distinguish different types of discrimination, some of them will be
briefly described. By protected attributes we mean those qualities,
traits or characteristics that one should not discriminate against.
Direct discrimination occurs when protected attributes of individuals
explicitly result in non-favorable outcomes toward them. Such
discrimination takes place for example when one uses the information
about someone's religious beliefs to evaluate whether they are competent
for the work they apply for. In contrast, in indirect discrimination
individuals appear to be treated equally, but anyway they end up being
treated unjustly due to the hidden effects of biases towards their
protected attributes. An example of such situation is when in a job
advert there is a requirement of 10 years experience instead of the list
of a specific type of experience and knowledge. This job advert can
discriminate indirectly against young people who can have the required
skills, but not yet so much work experience. Systemic discrimination
takes place when policies, customs or behaviors that result from certain
culture or organizational structure lead to discrimination against some
groups of people. Finally, statistical discrimination consists in
relying on group statistics to judge person belonging to a given group.
The topic of discrimination is entangled with another concept; fairness.
It is essential to grasp some concepts of fairness to take them into
consideration while designing implementation of some machine learning
model. In Mehrabi et al.
(\protect\hyperlink{ref-Mehrabi2019Survey}{2019}) one may notice that
depending on the context and application different definitions may be
applied.

There is a considerable amount of literature available on the topic of
bias detection and mitigation in NLP models. Bolukbasi, Chang, Zou,
Saligrama, \& Kalai (\protect\hyperlink{ref-Bolukbasi2016Man}{2016})
focus on gender biases that may be observable while investigating the
representation of job occupations and gender in terms of their assigned
numerical values. The authors apply cosine similarity measurement to
investigate the phenomenon where (the vectors corresponding to) words
related to jobs that are stereotypically associated with a given gender
are in fact in the model situated closer to this gender. They also use
analogy tasks to evaluate if bias is present in a word embedding model.
They check analogies by comparing pairs of word vectors. For example
they search for the word complementing the puzzle: man is to doctor as
woman is to \ldots{}? First they subtract the word ``man'' from the word
``woman''\footnote{This is pointwise subtraction of vectors, but for the
  simplicity we will often talk about words while really meaning
  vectors. What is meant should be clear from the context.} and then
they search for the ranked list of other words pairs that have similar
vector difference. They also include in the formula a threshold to
ensure that the resulting pairs could not be randomly picked.

However, as Nissim, Noord, \& Goot
(\protect\hyperlink{ref-Nissim2019Fair}{2019}) points out, there are
some limitations to this approach. In practice, most of analogies
implementations do not return any input words. This means that it does
not make sense to expect the algorithm to return the same profession for
both woman and man. Therefore, this method seems to be limited in terms
of bias detection. Other problems are related for example to the choice
of pairs and words that are used to detect the presence of
discrimination, as it is often subjective and without proper
justification. Additionally, the choice of the parameter used in
Bolukbasi et al. (\protect\hyperlink{ref-Bolukbasi2016Man}{2016})
formula to ensure that word pairs are not picked by random, is also not
justified and changing it drastically influences the results.

Islam, Bryson, \& Narayanan
(\protect\hyperlink{ref-Caliskan2017Semantics}{2016}) touch upon the
topic of biases regarding race and gender. They apply knowledge from
well-known psychological studies such as Implicit Association Test to
investigate the relation between human stereotypical thinking and
model-learnt biases to discover a close relationship between these two.
For the evaluation they use Word Embedding Association Test (WEAT) and
the Word Embedding Factual Association Test (WEFAT).

Manzini, Lim, Tsvetkov, \& Black
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}) propose a
novel way of using cosine similarity to verify if the assumed
resemblance between protected words and harmful attributes exists. They
investigate an approach that enables them to measure the bias for a
class (such as gender, religion or race) and express the final result
with a single metric. Their approach will be analyzed in details in the
next chapter.

Islam et al. (\protect\hyperlink{ref-Caliskan2017Semantics}{2016}) refer
to Implicit Association Test (Greenwald et al., 1998) that measures the
strength of associations between concepts or stereotypes by measuring
human reaction time for special tasks. Humans naturally exhibit some
biases which do not always cause social concern. One can imagine the
intuitive associations between for example insects and flowers, and the
feelings of pleasantness or unpleasantness. In general, people would
rather associate flowers with feeling pleasant than insects, and this
preference in a sense is a bias or prejudice in some direction. However,
this type of preference does not cause an uproar and is a rather morally
neutral case. This example is rather used to show that the methodology
used in IAT makes sense. Islam et al.
(\protect\hyperlink{ref-Caliskan2017Semantics}{2016}) were able to
obtain similar results to the ones that measured reaction time with the
use of WEAT metric. Unfortunately, further studies discover other biases
and prejudices that directly influence the quality of other people's
lives and therefore they should be taken care of.

The most common bias detection methods in the literature focus on
comparing the similarity between words from protected groups and those
that are considered to be stereotypical or harmful in some way. One can
find in this group methods such as euclidean distance or cosine
similarity (which is equivalent to dot product if the vectors are
normalized). There are also other ways to detect the effects of biases.
One can investigate the model performance on certain downstream tasks
and validate if the model uses the gender or race information to
complete the task. The currently used methods employing cosine
similarity, often aggregate the similarity values often aggregated in a
way that may lead to hasty conclusions. The averaging of values and the
lack of attention to uncertainty involved in such aggregation may lead
to the incomplete picture of the situation.

In the paper we indicate how current methods used to detect biases in
natural language models are limited, if we investigate them from the
perspective of Bayesian analysis. Our research enhances the current way
in which the bias detection is performed to make sure that it is
methodologically valid. The key hypothesis is that greater understanding
of data and bias implications can be achieved when Bayesian methods are
applied to issue.

\chapter{Cosine similarity and bias
detection}\label{cosine-similarity-and-bias-detection}

\section{Word embeddings}\label{word-embeddings}

To understand what cosine similarity measurement is, one first needs to
grasp the concept of translating words to a computer-readable form. In
the field of natural language processsing there are two main types of
words representation --- localist and distributed. One-hot encoding is
an example of a method used to achieve a localist representation of
words. Here, each vector contains information only about a single data
point. This is achieved by first mapping categorical values (words) to
integers and then to each integers a binary vector is assigned which
contains only 0s except for the index of the integer, which is assigned
1. An example of a localist representation is:

\begin{longtable}[]{@{}llllll@{}}
\toprule
word & 1 & 2 & 3 & 4 & 5\tabularnewline
\midrule
\endhead
woman & 1 & 0 & 0 & 0 & 0\tabularnewline
man & 0 & 1 & 0 & 0 & 0\tabularnewline
girl & 0 & 0 & 1 & 0 & 0\tabularnewline
boy & 0 & 0 & 0 & 1 & 0\tabularnewline
monarch & 0 & 0 & 0 & 0 & 1\tabularnewline
\bottomrule
\end{longtable}

In the example above it is clear that the length of the vectors
increases with the number of words in a vocabulary. It is not a very
computationally efficient representation. It has other flaws as well.
For example, it is unable to capture the resemblance between words
appearing in similar contexts.

In contrast to the localist representation, a distributed representation
returns vectors that contain continuous values instead of discrete 1s
and 0s. Word embeddings are a class of various techniques that allow one
to represent words as distributed vectors. Such learned representations
of text have certain properties. At least prima facie, they store
similar (or at least co-occurring) words close to each other in a vector
space. An example of distributed representation is:

\begin{longtable}[]{@{}lllll@{}}
\toprule
word & 1 & 2 & 3 & 4\tabularnewline
\midrule
\endhead
woman & 0.456 & 0.267 & 0.675 & 0.131\tabularnewline
man & 0.451 & 0.897 & 0.472 & 0.088\tabularnewline
girl & 0.604 & 0.262 & 0.414 & 0.706\tabularnewline
boy & 0.279 & 0.172 & 0.475 & 0.010\tabularnewline
monarch & 0.565 & 0.678 & 0.463 & 0.975\tabularnewline
\bottomrule
\end{longtable}

One of the advantages of using a distributed representation is that one
is able to represent an enormous number of concepts with a smaller
number of units. It is also possible to better capture similarities as
words of similar meanings can have similar numeric vectors.

The numbers occurring in such representations are not random. They are
learned in a process that uses a very shallow neural network. There are
various types of techniques used for learning the vectors
representations. One of the most straightforward ones is a skip-gram
model. Given a word the model tries to predict its neighboring words
from the sentence. The mathematics behind the process relies on the idea
that the prediction concerns the conditional probability of the adjacent
words. The algorithm tries to minimize the loss function, which
penalizes the system for discrepancy with actual co-occurrence
frequencies in the corpus. One can choose various parameters of the
model, such as the window size that determines how many surrounding
words the model should predict. After preparing such a fitted model, one
takes only the learned weights from a neural network, and uses them as
vectors in a word embeddings representation.

Word embeddings have many applications in natural language processing.
They are handy in document search and information retrieval. They also
play their part in improving automatic translations. Well learned word
representations may also contribute to the improvement of sentiment
analysis or spam detection.

\section{Cosine similarity and
distance}\label{cosine-similarity-and-distance}

Cosine similarity is often used as a method of finding out whether
vector representations for two words suggest that they are similar or
somehow connected. Cosine similarity is the cosine of the angle between
two vectors: the result of dividing their inner product (dot product
usually) by the product of their magnitudes.

It is worth mentioning one more point concerning cosine similarity.
After the vectors are normalized to have length equal to 1, the inner
product itself (often dot product) is used to measure the similarity,
because it then equals the cosine similarity.

\begin{align} \tag{Sim}
\mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\vert \vert A \vert \vert \,\vert \vert B \vert \vert}
\end{align}

Cosine similarity is considered a proper tool for this operation as its
result has a clear connection to geometry and at least for a low number
of dimensions may be easily interpreted. Using this scale, one can
compare vector similarities in a fairly clear manner. When the vectors
are aligned perpendicularly to each other, their similarity equals 0
(which is the same as the cosine of 90 degrees). This tells us that the
similarity between the vectors is small. As the angle between vectors
decreases, cosine similarity approaches one, which stands for the
greatest similarity. It is also possible to obtain negative cosine
similarity, if the value approaches -1 it means that the words are
contrary to each other.

One of the limitations of this measure is that it informs us only about
similarities between vectors in terms of their orientation. However, it
is often argued that in comparing words in terms of this metric, the
magnitude of vectors may be treated as irrelevant, as the most important
information pertains to direction.

In what follows, it is important to distinguish between cosine
similarity and cosine distance, defined as:

\begin{align} \tag{Sim}
\mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B)\\
 &  = 1 - \frac{A \cdot B}{\vert \vert A \vert \vert \,\vert \vert B \vert \vert} \nonumber
\end{align}

The greater the similarity between two vectors, the smaller the distance
between them. The cosine distance ranges between 0 and 2. If the vectors
are in an opposite directions, the cosine distance is 2. And if the
vectors are extremely similar then the cosine distance is very close to
0.

One should note that cosine distance is not exactly a distance measure,
as it does not meet triangle inequality requirements. The triangle
inequality formula says that for any triangle, the sum of the lengths of
any two sides must be greater than or equal to the length of the
remaining side. As shown in
\href{https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u}{discussion
in stats.stackexchange.com} in the case of cosine distance it would have
to fulfill this equation 1+cos-sim(𝐴,𝐶) \textless{} cos-sim(𝐴,𝐵) +
cos-sim(𝐵,𝐶). If one chooses specific unit vectors it is easy to
demonstrate that the triangle inequality is not preserved.

\section{Cosine distance in a one-class bias
detection}\label{cosine-distance-in-a-one-class-bias-detection}

Bolukbasi et al. (\protect\hyperlink{ref-Bolukbasi2016Man}{2016}) define
similarity between words as the outcome inner product of their
normalized vectors. They focus on examining what the geometry of word
embedding is in regard to ``he'' and ``she'' words. In other words,
whether the similarity between those concepts and other words reflects
expected gender stereotypes. They test this hypothesis by investigating
whether there is a connection between word embeddings representing
certain professions and words referring to gender. They also evaluate
whether automatically produced analogies between words reflect the
stereotypes as well.

Here the gender bias of a word \(w\) is understood as its projection on
the gender direction
\(\vec{w} \cdot (\overrightarrow{he} - \overrightarrow{she})\) (the
gender direction is the top principal compontent of ten gender pair
difference vectors). The underlying idea is that no bias is present if
non-explicitly gender words are in equal distance to both elements in
all explicitly gender pairs. Given the (ideally) gender neutral words
\(N\) and the gender direction \(g\) the direct direct gender bias is
defined as the average distance of the words in \(N\) from \(g\) (\(c\)
is a parameter determining how strict we want to be):

\begin{align}
\mathsf{directBias_c(N,g)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},g)\vert^c}{\vert N \vert }
\end{align}

A very vivid way to follow their method of arguing that bias in word
embeddings is real is to plot the values of inner product of chosen
words. The plot below does not originate from the original paper but
from
\href{https://www.kaggle.com/rtatman/gender-bias-in-word-embeddings}{kaggle.com
R notebook using data from GloVe: Global Vectors for Word
Representation}. However, similar visualization may be found there. Data
used to create our plot is as follows.

Occupations associated with feminine:
\textbf{"homemaker", "nurse", "receptionist", "librarian", "socialite", "hairdresser", "nanny", "bookkeeper", "stylist", "housekeeper", "interior designer", "guidance counselor"}

Occupations associated with masculine:
\textbf{"maestro", "skipper", "protege", "philosopher", "captain", "architect", "financier", "warrior", "broadcaster", "magician", "fighter pilot", "boss"}

\vspace{1mm} \footnotesize

\begin{center}\includegraphics[width=1\linewidth]{_main_files/figure-latex/unnamed-chunk-1-1} \end{center}

\normalsize

The points in the plot above result from the calculation of the inner
product of a chosen vector for a profession word and a vector for a
gender word (she or he). Inner product of two vectors expresses
similarity between words. This assumption originates from the geometry
and properties of a vector space. In the picture one may observe a
correlation between the occupation and gender. Stereotypical male
professions are closer to the ``he'' word and stereotypical female
professions have greater similarity with ``she'' word. One could argue
that this is a hint that there is some hidden bias information stored in
the words vectors.

\section{Cosine distance in a multi-class bias
detection}\label{cosine-distance-in-a-multi-class-bias-detection}

Manzini et al.
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}) present a
different approach towards finding similarities between classes of
words. The authors claim that texts available online are full of direct
or indirect human stereotypes. As a result, word embeddings are prone to
learn and maybe amplify those biases and propagate them further into AI
models that are used for various applications. Cosine distance is used
in the article as a measure to first argue for the existence of
multi-class bias and then to show how through bias mitigation techniques
the bias may be decreased.

They modify WEAT to a multi-class setting, introducing Mean Average
Cosine similarity as a measure of bias (in fact, in the paper they
report distances rather than similarities). Let
\(T = \{t_1, \dots, t_k\}\) be a class of protected word embeddings, and
let each \(A_j\in A\) be a set of attributes stereotypically associated
with a protected word). Then:

\begin{align}
S(t_i, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) \\
MAC(T,A) & = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j)
\end{align}

As the code has been provided by the authors we were able to reconstruct
their results. The main steps in the procedure are as follows. Let us go
through an example that refers to the process of hard debiasing on
religious attributes.

First we load word embeddings from
reddit.US.txt.tok.clean.cleanedforw2v.w2v dataset. The word embeddings
have only 50 dimensions and the number of individual words from the
dataset is 44895. The authors assume that the protected group should
ideally not have high cosine similarity to stereotypical words. The word
embeddings geometry should not place this group close to harmful
stereotypes, if it is to be bias-free. For instance, let's look at the
religion-related words.

Protected words by religion type:

\textbf{"jew"       : ["judaism", "jew", "synagogue", "torah", "rabbi"]}

\textbf{"christian" : ["christianity", "christian", "church", "bible", "priest"]}

\textbf{"muslim"    : ["islam", "muslim", "mosque", "quran", "imam"]}

Stereotypical words by religion type:

\textbf{"jew"       : ["greedy", "cheap", "hairy", "liberal"]}

\textbf{"christian"       : ["judgemental", "conservative", "familial"]}

\textbf{"muslim"       : ["violent", "terrorist", "dirty", "uneducated"]}

We have prepared a table presenting the values of cosine distance for
each protected word with each attribute (stereotype). The part of the
results is shown in Table 2.3

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{religion <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"../datasets/religionReddit.csv"}\NormalTok{)[}\OperatorTok{-}\DecValTok{1}\NormalTok{]}
\KeywordTok{colnames}\NormalTok{(religion) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"protectedWord"}\NormalTok{,}\StringTok{"wordToCompare"}\NormalTok{,}\StringTok{"wordClass"}\NormalTok{,}
                        \StringTok{"cosineDistance"}\NormalTok{,}\StringTok{"cosineSimilarity"}\NormalTok{,}\StringTok{"connection"}\NormalTok{)}
\NormalTok{religion}\OperatorTok{$}\NormalTok{wordClass <-}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(religion}\OperatorTok{$}\NormalTok{wordClass)}
\KeywordTok{levels}\NormalTok{(religion}\OperatorTok{$}\NormalTok{wordClass) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"christian"}\NormalTok{,}\StringTok{"human"}\NormalTok{,}\StringTok{"jewish"}\NormalTok{,}\StringTok{"muslim"}\NormalTok{,}\StringTok{"neutral"}\NormalTok{)}
\KeywordTok{head}\NormalTok{(religion)  }\OperatorTok{%>%}\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{format =} \StringTok{"latex"}\NormalTok{,}\DataTypeTok{booktabs=}\NormalTok{T,}
                      \DataTypeTok{linesep =} \StringTok{""}\NormalTok{,  }\DataTypeTok{escape =} \OtherTok{FALSE}\NormalTok{, }
                      \DataTypeTok{caption =} \StringTok{"Head of the religion dataset."}\NormalTok{) }\OperatorTok{%>%}
\StringTok{                      }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options=}\KeywordTok{c}\NormalTok{(}\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:religionTableHeadEarly}Head of the religion dataset.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllrrl}
\toprule
protectedWord & wordToCompare & wordClass & cosineDistance & cosineSimilarity & connection\\
\midrule
judaism & violent & muslim & 0.7141939 & 0.2858061 & different\\
judaism & terrorist & muslim & 0.7461333 & 0.2538667 & different\\
judaism & dirty & muslim & 1.2002599 & -0.2002599 & different\\
judaism & uneducated & muslim & 0.7885469 & 0.2114531 & different\\
judaism & greedy & jewish & 1.0026172 & -0.0026172 & associated\\
judaism & cheap & jewish & 1.2323229 & -0.2323229 & associated\\
\bottomrule
\end{tabular}}
\end{table}

\normalsize

\pagebreak

In the article there was no analysis of individual distances, but the
general look at the means. The authors introduced a metric that tries to
generalize WEAT in measuring the presence of bias through the
classification of multi-class bias in groups of words connected with
gender, religion or race. In the process they first take the mean of
cosine distances between a given protected word and attributes assigned
to each stereotype. They do not differentiate between stereotypes
associated with a word and stereotypes associated with different words
(in the case of religion, stereotypes characteristic for Christianity
has also cosine distance measured with for instance, Judaism or Islam).
Then, after collecting the list of mean cosine distances, they average
the list to obtain one final value representing the whole group, in this
example religion, for which the final mean of all mean distances is
equal to 0.859.

In the article the authors also try to remove biases, as previously
defined, from a word embedding. First they identify the bias subspace
using Principal Component Analysis (PCA) which is a technique for
dimensionality reduction. It is applied here to choose the subspace that
contains the greatest amount of information. There can be many subspaces
found in a given group. For example in terms of religion one can
identify at least a few sets that are to grasp the concept of religion
in general:

\textbf{["judaism", "christianity", "islam"]}

\textbf{["jew", "christian", "muslim"]}

\textbf{["synagogue", "church", "mosque"]}

The idea is to find a set that provides enough information to create
from it a vector representing the concept of religion among words. This
strategy is based on the idea that different dimensions of vectors
contain different types of information and in some words in vector
layers (subspaces) the information about religiousness is implicitly
conveyed. In some cases this knowledge is useful, but in the case of
harmful stereotypes one does not want to include the concept of religion
in stereotypical words.

After finding the bias subspace, they use it to modify the vector values
individually so that their cosine distances towards certain words are
changed. In the case of stereotypes the aim is to make the cosine
distances larger so that the association between protected word and
stereotype is smaller.

In the final step there is an evaluation of the results. The cosine
distances are calculated again but this time using debiased vocabulary.
After taking the mean of all distances one final value is obtained and
then it is compared with the average value from the beginning. If the
cosine distances are on average greater than before then it leads the
authors to the conclusion that improvement has been achieved. As the
cosine distance increases, it is assumed that the association between
protected and stereotypical words decreases.

\section{Limitations of the approach}\label{limitations-of-the-approach}

\textbf{1. Selection of attributes}

The attributes are taken from different sources, there is no principled
justification for their choice. From our analysis it will become clear
that the list is rather uneven.

There is no mention of methodology for deciding on the number of
attributes necessary to decide a hypothesis on the given size of
dataset. There are however some ways to estimate sample sizes needed to
ensure that the results are meaningful if the effect is present. Our
research will show that the numbers used are rather insufficient.

\textbf{2. No attention to distributions and details}

The authors use the mean of mean average cosine similarities to measure
multi-class similarity between protected word and harmful stereotypes.
They average the results until they obtain one final value to represent
the mean cosine distance between protected word from a given class and
the attributes of that class. If one takes closer look at the individual
values that are taken for the calculations it turns out that there is a
bunch of outliers and surprisingly dissimilar words. We approach this
issue by providing the tables and visualizations of individual cosine
distances to make sure that we get the proper insight into the data.

\textbf{3. Hiding the impact of uncertainty}

With such a method the uncertainty is not included which makes it even
more difficult to give reasonable interpretations of the results. We
propose the use of Bayesian method to obtain some understanding of the
influence the uncertainty has on the interpretation of final results.

\textbf{4. No word class distinction}

In the original paper, words from all three religions were compared
against all of the stereotypes, which means that there was no
distinction between cases in which the stereotype is associated with a
given religion, as opposed to the situation in which it is associated
with another one. Not all of the stereotypical words have to be
considered as harmful for all of the religions. In our analysis, we
distinguish between stereotypes associated with a given group,
stereotypes associated with different groups, and control groups:
neutral words and stereotypes-free human predicates.

\textbf{5. Interpreting the results}

Assuming for a moment that the value of multi-class cosine distance is
correct, one may question the interpretation. Manzini et al.
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}) summarize the
averages of cosine distance per group (gender, race, religion). For now
let us focus now on analyzing the values relating to religious biases.
Here is the relevant fragment of table:

\begin{longtable}[]{@{}ll@{}}
\toprule
Religion Debiasing & MAC\tabularnewline
\midrule
\endhead
Biased & 0.859\tabularnewline
Hard Debiased & 0.934\tabularnewline
Soft Debiased (\(\lambda\) = 0.2) & 0.894\tabularnewline
\bottomrule
\end{longtable}

MAC stand for mean average cosine similarity, although in reality the
the table contains mean distances. What may attract attention is the
fact that the value of cosine distance in ``Biased'' category is already
quite high even before debiasing. High cosine distance indicates low
cosine similarity between values. One could think that average cosine
similarity equal to approximately 0.141 is not significant enough to
consider it as bias. However the authors aim to mitigate ``biases'' in
vectors with such great distance to make it even larger.
Methodologically the question is, on what basis is this small similarity
still considered as a proof of the presence of bias, and whether these
small changes are meaningful. This is in general the problem of scale
and the lack of universal intervals. In contrast, proper intervals
decide whether given cosine similarity is high enough to consider the
words to be more similar than if we chose them at random. We will use
highest posterior density intervals.

\textbf{6. The curse of dimensionality}

In our case, the curse of dimensionality may take place when there is an
increase in the volume of data that results in adding extra dimensions
to the Euclidean space. According to the article
\href{https://analyticsindiamag.com/curse-of-dimensionality-and-what-beginners-should-do-to-overcome-it/}{Curse
of dimensionality at analyticsindiamag.com} as the number of features
increases, it may be harder and harder to obtain useful information from
the data using the available algorithms. One may notice that more data
should contribute to greater amount of information, but more information
also means greater risk of noise and distractions in data. At the same
time, modern solutions are often adapted to smaller dimensions and their
results in higher ones are not intuitive, or may be prone to error.

Using cosine similarity in high dimensions in word embeddings may also
be prone to the curse of dimensionality. According to Venkat
(\protect\hyperlink{ref-Venkat2018Curse}{2018}) there are reasons to
consider this phenomenon when searching for word similarities in higher
dimensions. An experiment is conducted that aims at showing how the
similarity values and variation change as the number of dimensions
increases. The hypothesis made in the paper states that two things will
happen as the number of dimensions increase. First, the effort required
to measure cosine similarity will be greater, and two, the similarity
between data will blur out and have less variation. The authors generate
random points with increasing number of dimensions where each dimension
of a data point is given a value between 0 and 1. Then they pick one
vector at random from each dimension class and calculate the cosine
similarity between the chosen vector and the rest of the data. Then they
check how the variation of values changes as the number of dimensions
increases. It seems like the more dimensions there are, the smaller the
variance and therefore it is less obvious how to interpret the resulting
cosine similarities. Maybe the scale should be adjusted to the number of
dimensions and variance so that it still gives us sensible information
about data. Yet, according to some articles cosine similarity in high
dimensions is not reliable enough as it may be the case that choosing
words at random may result in similar values as when picking them
consciously.

\begin{figure}
\centering
\includegraphics{../images/curseOfDimensionality.png}
\caption{curse of dimensionality, number of dimensions on the x axis,
standard deviation of similarity on the y axis}
\end{figure}

\chapter{Walkthrough with the religion
dataset}\label{walkthrough-with-the-religion-dataset}

\section{Loading and understanding the
dataset}\label{loading-and-understanding-the-dataset}

We start with loading the libraries needed for the analysis.

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(ggthemes)}
\KeywordTok{library}\NormalTok{(rethinking)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(ggpubr)}
\KeywordTok{library}\NormalTok{(kableExtra)}
\KeywordTok{library}\NormalTok{(dplyr)}
\KeywordTok{library}\NormalTok{(ggExtra)}
\KeywordTok{library}\NormalTok{(cowplot)}
\end{Highlighting}
\end{Shaded}

\normalsize

We will use the choice of protected words and stereotypical predicates
used in REF. This is a decent point of departure, not only we want to
compare our method to that of REF, but also because this data format is
fairly general (as contrasted, say, with a set up for binary
stereotypes). Note also that the method we develop here can fairly
easily be run for different stereotypization patterns. Let's start with
explaining the method and its deployment using a dataset obtained for
the religion-related protected words.

Let's load, clean a bit and inspect the head of the religion dataset we
prepared. In order to obtain this dataset, we calculated the cosine
distance between each protected word and each word from both the
bias-related attribute groups, which were used in the original study,
and to neutral and human control attributes which we added as control
groups. For instance, for religion, the bias-related predicates (coming
from the original study in REF) include muslim bias attributes, jew bias
attributes, christian bias attributes (see a list in the APPENDIX).

We decided to add control groups in the form of two classes --- neutral
words and human-related words. Without a proper control group it is
quite hard to compare the resulting cosine distances and decide on their
significance in bias detection. We prepared approximately 230 neutral
words to double-check the prima-facie neutral hypothesis that their
cosine similarity to the protected words will oscillate around 0 (that
is, the distances will be around 1). This provides us with a more
reliable point of reference. Moreover, we added human attributes that
are associated with people in general to investigate whether the smaller
cosine distance between protected words and stereotypes can result
simply from the fact that the stereotype predicates are associated with
humans. For two control groups, we have randomly drawn 230 words that do
not express any property usually attributed to humans, and human related
attributes.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{religion <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"../datasets/religionReddit.csv"}\NormalTok{)[}\OperatorTok{-}\DecValTok{1}\NormalTok{]}
\KeywordTok{colnames}\NormalTok{(religion) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"protectedWord"}\NormalTok{,}\StringTok{"wordToCompare"}\NormalTok{,}\StringTok{"wordClass"}\NormalTok{,}
                        \StringTok{"cosineDistance"}\NormalTok{,}\StringTok{"cosineSimilarity"}\NormalTok{,}\StringTok{"connection"}\NormalTok{)}
\CommentTok{#head(religion)}
\CommentTok{#library(plyr)}
\NormalTok{religion}\OperatorTok{$}\NormalTok{wordClass <-}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(religion}\OperatorTok{$}\NormalTok{wordClass)}
\CommentTok{#levels(religion$wordClass)}
\CommentTok{#religion$wordClass <- mapvalues(religion$wordClass, }
\CommentTok{#                          from = levels(religion$wordClass),}
\CommentTok{#                          to = c("christian","human","jewish","muslim","neutral"))}
\KeywordTok{levels}\NormalTok{(religion}\OperatorTok{$}\NormalTok{wordClass) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"christian"}\NormalTok{,}\StringTok{"human"}\NormalTok{,}\StringTok{"jewish"}\NormalTok{,}\StringTok{"muslim"}\NormalTok{,}\StringTok{"neutral"}\NormalTok{)}
\KeywordTok{head}\NormalTok{(religion)  }\OperatorTok{%>%}\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{format =} \StringTok{"latex"}\NormalTok{,}\DataTypeTok{booktabs=}\NormalTok{T,}
                      \DataTypeTok{linesep =} \StringTok{""}\NormalTok{,  }\DataTypeTok{escape =} \OtherTok{FALSE}\NormalTok{, }
                      \DataTypeTok{caption =} \StringTok{"Head of the religion dataset."}\NormalTok{) }\OperatorTok{%>%}\StringTok{      }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{latex_options=}\KeywordTok{c}\NormalTok{(}\StringTok{"scale_down"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:religionTableHead}Head of the religion dataset.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllrrl}
\toprule
protectedWord & wordToCompare & wordClass & cosineDistance & cosineSimilarity & connection\\
\midrule
judaism & violent & muslim & 0.7141939 & 0.2858061 & different\\
judaism & terrorist & muslim & 0.7461333 & 0.2538667 & different\\
judaism & dirty & muslim & 1.2002599 & -0.2002599 & different\\
judaism & uneducated & muslim & 0.7885469 & 0.2114531 & different\\
judaism & greedy & jewish & 1.0026172 & -0.0026172 & associated\\
judaism & cheap & jewish & 1.2323229 & -0.2323229 & associated\\
\bottomrule
\end{tabular}}
\end{table}

\normalsize

The \texttt{protectedWord} column contains words from a protected class
that (in a perfect world according to the assumptions of the orignal
study) should not be associated with harmful stereotypes.
\texttt{wordToCompare} contains attributes, including stereotypes and
control group words. For each row we compute the cosine distances
between a given protected word and a given attribute word.
\texttt{wordClass} tells us which class an attribute is supposed to be
stereotypically associated with, that is, whether the word from
\texttt{wordToCompare} is associated stereotypically with jews,
christians or muslims, or whether it belongs to a control group.
\texttt{cosineDistance} is simply a calculation of the cosine distance
between protected word and atrribute. \texttt{cosineSimilarity} contains
the result of substracting cosine distance from 1. \texttt{connection}
contains information about the relation type between a protected word
and an attribute. If the attribute is e.g.~a harmful jewish stereotype
and the protected word is also from the judaism group, the connection
has value \texttt{associated}. If the attribute is still stereotypically
jewish, but the protected word comes from another religion, the
connection is labelled as \texttt{different}. If the attribute belongs
to a neutral group then the connection is labelled as \texttt{none} and
if an attribute belongs to the \texttt{human} class, then the connection
is labelled as \texttt{human}.

\section{First look at the empirical
distributions}\label{first-look-at-the-empirical-distributions}

First let's take a look at the empirical distribution of distances by
the connection type, initially ignoring the human control class for now.

\vspace{1mm} \footnotesize

\begin{center}\includegraphics[width=1\linewidth]{_main_files/figure-latex/unnamed-chunk-3-1} \end{center}

\normalsize

The first impression is that while there is a shift for associated words
towards smaller cosine distances as compared to the neutral words,
slightly surprisingly a slightly weaker shift in the same direction is
visible for attributes associated with different stereotypes. Moreover,
the empirical distributions overlap to alarge extent and the means
grouped by connection type do not seem too far from each other. In fact,
as there is a lot of variety in the cosine distances (as we will soon
see), we need to gauge the uncertainty involved, and to look more
carefully at individual protected words to get a better idea of how the
cosine distance distribution changes for different attribute groups and
different protected classes. Now, let's add the human attributes to the
picture:

\vspace{1mm} \footnotesize

\begin{center}\includegraphics[width=1\linewidth]{_main_files/figure-latex/unnamed-chunk-4-1} \end{center}

\normalsize

\noindent Notice that the distribution for \texttt{human} (even though
we did our best not to include in it any stereotype-related atributes)
is left-skewed, with much overlap with \texttt{associated} and
\texttt{different}, which illustrates the need to take being associated
with humans as an important predictor.

Our focus lies in \texttt{connection} as a predictor. Morever, later on
we'll be interested in looking at the protected words separately, and at
protected words split by connection. For technical reasons it is useful
to represent these factors as integer vectors.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{religion}\OperatorTok{$}\NormalTok{con <-}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(religion}\OperatorTok{$}\NormalTok{connection)}
\NormalTok{religion}\OperatorTok{$}\NormalTok{pw <-}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(religion}\OperatorTok{$}\NormalTok{protectedWord)}
\NormalTok{religion}\OperatorTok{$}\NormalTok{pwFactor <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(religion}\OperatorTok{$}\NormalTok{protectedWord, religion}\OperatorTok{$}\NormalTok{connection))}
\NormalTok{religion}\OperatorTok{$}\NormalTok{pwIndex <-}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(religion}\OperatorTok{$}\NormalTok{pwFactor)}
\end{Highlighting}
\end{Shaded}

\normalsize

A short script, \texttt{cleanDataset} to make this faster, so
equivalently:

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{source}\NormalTok{(}\StringTok{"../functions/cleanDataset.R"}\NormalTok{)}
\NormalTok{religion <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"../datasets/religionReddit.csv"}\NormalTok{)[}\OperatorTok{-}\DecValTok{1}\NormalTok{]}
\NormalTok{religion <-}\StringTok{ }\KeywordTok{cleanDataset}\NormalTok{(religion,}\KeywordTok{c}\NormalTok{(}\StringTok{"christian"}\NormalTok{,}\StringTok{"human"}\NormalTok{,}\StringTok{"jewish"}\NormalTok{,}\StringTok{"muslim"}\NormalTok{,}\StringTok{"neutral"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\normalsize

\section{Looking at the islam-related
words}\label{looking-at-the-islam-related-words}

For now, let's focus on five protected words related to islam (``imam'',
``islam'', ``mosque'', ``muslim'', and ``quran''). The word list
associates with islam four stereotypical attributes (``violent'',
``terrorist'', ``uneducated'' and ``dirty''). First, we select and plot
the empirical distributions for these protected words.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\NormalTok{muslimWords <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"imam"}\NormalTok{,}\StringTok{"islam"}\NormalTok{,}\StringTok{"mosque"}\NormalTok{,}\StringTok{"muslim"}\NormalTok{,}\StringTok{"quran"}\NormalTok{)}
\NormalTok{muslim <-}\StringTok{ }\NormalTok{religion }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(protectedWord }\OperatorTok{%in%}\StringTok{ }\NormalTok{muslimWords)}
\KeywordTok{ggplot}\NormalTok{(muslim, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{  cosineDistance, }\DataTypeTok{fill =}\NormalTok{ connection, }\DataTypeTok{color =}\NormalTok{ connection))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_density}\NormalTok{(}\DataTypeTok{alpha=}\FloatTok{0.6}\NormalTok{,}\DataTypeTok{size =} \FloatTok{.2}\NormalTok{)}\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"orangered4"}\NormalTok{,}\StringTok{"chartreuse4"}\NormalTok{, }\StringTok{"skyblue"}\NormalTok{, }\StringTok{"gray"}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{breaks =} \KeywordTok{seq}\NormalTok{(}\FloatTok{0.3}\NormalTok{,}\FloatTok{1.5}\NormalTok{, }\DataTypeTok{by =} \FloatTok{0.1}\NormalTok{))}\OperatorTok{+}\KeywordTok{xlab}\NormalTok{(}\StringTok{"cosine distance"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"orangered4"}\NormalTok{,}\StringTok{"chartreuse4"}\NormalTok{,}\StringTok{"skyblue"}\NormalTok{,}\StringTok{"gray"}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_tufte}\NormalTok{()}\OperatorTok{+}\KeywordTok{ggtitle}\NormalTok{(}\StringTok{"Empirical distribution of distances (muslim)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{_main_files/figure-latex/unnamed-chunk-7-1} \end{center}

\normalsize

\noindent Once we focus on words related to islam, the associated bias
seems to be stronger than in the whole dataset. This is a step towards
illustrating that the distribution of bias is uneven.

Now, say we want to look at a single protected word. Since the dataset
also contains comparison multiple control neutral and human attributes,
we randomly select only 5 from \texttt{none} and 5 from \texttt{human}
control groups of those for the visualisation purposes.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\NormalTok{muslimClass <-}\StringTok{ }\NormalTok{muslim }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(protectedWord }\OperatorTok{==}\StringTok{ "muslim"}\NormalTok{)}
\NormalTok{neutralSample <-}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\KeywordTok{filter}\NormalTok{(muslimClass,connection }\OperatorTok{==}\StringTok{ "none"}\NormalTok{), }\DecValTok{5}\NormalTok{)}
\NormalTok{humanSample <-}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\KeywordTok{filter}\NormalTok{(muslimClass,connection }\OperatorTok{==}\StringTok{ "human"}\NormalTok{), }\DecValTok{5}\NormalTok{)}
\NormalTok{muslimVis <-}\StringTok{ }\NormalTok{muslimClass }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(connection }\OperatorTok{!=}\StringTok{ "none"} \OperatorTok{&}\StringTok{ }\NormalTok{connection }\OperatorTok{!=}\StringTok{"human"}\NormalTok{)}
\NormalTok{muslimVis <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(muslimVis,neutralSample,humanSample)}

\CommentTok{#we plug in our visualisation script}
\KeywordTok{source}\NormalTok{(}\StringTok{"../functions/visualisationTools.R"}\NormalTok{)}
\CommentTok{#two arguments: dataset and protected word}
\KeywordTok{visualiseProtected}\NormalTok{(muslimVis,}\StringTok{"muslim"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{_main_files/figure-latex/tableMuslimActive-1} \end{center}

\normalsize

Note that the distance between the grey point and the other points is
proportional to cosine distance, the non-grey point size is proportional
to cosine similarity to the protected word, and color groups by the
connection type. So for \texttt{muslim} it seems that the stereotypes
coming from the word list are fairly well visible. To give you some
taste of how uneven the dataset is, compare this to what happens with
\texttt{priest}.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\NormalTok{priestClass <-}\StringTok{ }\NormalTok{religion }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(protectedWord }\OperatorTok{==}\StringTok{ "priest"}\NormalTok{)}
\NormalTok{neutralSample <-}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\KeywordTok{filter}\NormalTok{(priestClass,connection }\OperatorTok{==}\StringTok{ "none"}\NormalTok{), }\DecValTok{5}\NormalTok{)}
\NormalTok{humanSample <-}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\KeywordTok{filter}\NormalTok{(priestClass,connection }\OperatorTok{==}\StringTok{ "human"}\NormalTok{), }\DecValTok{5}\NormalTok{)}
\NormalTok{priestVis <-}\StringTok{ }\NormalTok{priestClass }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(connection }\OperatorTok{!=}\StringTok{ "none"} \OperatorTok{&}\StringTok{ }\NormalTok{connection }\OperatorTok{!=}\StringTok{"human"}\NormalTok{)}
\NormalTok{priestVis <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(priestVis,neutralSample,humanSample)}

\CommentTok{#we plug in our visualisation script}
\KeywordTok{source}\NormalTok{(}\StringTok{"../functions/visualisationTools.R"}\NormalTok{)}
\CommentTok{#two arguments: dataset and protected word}
\KeywordTok{visualiseProtected}\NormalTok{(priestVis,}\StringTok{"priest"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{_main_files/figure-latex/unnamed-chunk-8-1} \end{center}

\normalsize

\noindent Here you can see that some human attributes are closer than
stereotype attributes, and that there is no clear reason to claim that
\texttt{associated} attributes are closer than \texttt{different} or
\texttt{human} attributes. This, again, illustrates the need of
case-by-case analysis with control groups.

The general idea now is that the word lists provided in different pieces
of research are just samples of attributes associates with various
stereotypes and should be treated as such: the uncertainty involved and
the sample sizes should have clear impact on our estimates.

\section{Bayesian model structure and
assumptions}\label{bayesian-model-structure-and-assumptions}

We will now think of cosine distance as the output variable, and will
build a few Bayesian models to compare. First, we just build a baseline
model which estimates cosine distance to the attributes separately for
each protected word. The underlying idea is that different protected
words might in general have different relations to all the attributes
and these relations should be our point of departure.

Here is the intuition behind the mathematical Bayesian model involved.
Our outcome variable is \texttt{cosine\ difference}, which we take to me
normally distributed around the predicted mean for a given protected
word (that is, we assume the residuals are normally distributed). The
simplest model specification is:

\begin{align}
cosineDistance_i  & \sim dnorm(\mu_i, \sigma) \\
\mu_i & = m_{pw} \\
m_{pw} & ~ dnorm(1,.5) \\
\sigma &\sim  dcauchy(0,1)
\end{align}

That is, we assume the estimated means might be different for diferent
protected words and our prior for the mean and the overal standard
deviation are normal with mean 1 and sd=.5 and half-cauchy with
parameters \texttt{0,1}. Further on we'll also estimate additional
impact the connection type may have. For this impact we take a slightly
skeptical prior centered around 0 distributed normally with sd = 1.
These are fairly weak and slightly skeptical regularizing priors, which
can be illustrated as follows:

\vspace{2mm}

\begin{center}\includegraphics[width=1\linewidth]{_main_files/figure-latex/priorsVis-1} \end{center}

\section{Choosing predictors}\label{choosing-predictors}

Now we can define and compile the baseline model. Its parameters will
have a posterior distribution obtained using either Hamiltionian Monte
Carlo methods (STAN) available through the \texttt{rethinking} package.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(rethinking)}
\KeywordTok{options}\NormalTok{(}\DataTypeTok{buildtools.check =} \ControlFlowTok{function}\NormalTok{(action) }\OtherTok{TRUE}\NormalTok{ )}
\NormalTok{religionBaseline <-}\StringTok{ }\KeywordTok{ulam}\NormalTok{(}
  \KeywordTok{alist}\NormalTok{(}
\NormalTok{    cosineDistance }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(mu,sigma),}
\NormalTok{    mu <-}\StringTok{ }\NormalTok{m[pw],}
\NormalTok{    m[pw] }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    sigma }\OperatorTok{~}\StringTok{ }\KeywordTok{dcauchy}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)}
\NormalTok{  ),}
  \DataTypeTok{data =}\NormalTok{ religion,}
  \DataTypeTok{chains=}\DecValTok{2}\NormalTok{ , }\DataTypeTok{iter=}\DecValTok{4000}\NormalTok{ , }\DataTypeTok{warmup=}\DecValTok{1000}\NormalTok{,}
  \DataTypeTok{start=} \KeywordTok{list}\NormalTok{(}\DataTypeTok{mu =} \DecValTok{1}\NormalTok{, }\DataTypeTok{co =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sigma=} \FloatTok{.3}\NormalTok{),}
  \DataTypeTok{log_lik =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{cores=}\DecValTok{4}
\NormalTok{)}
\CommentTok{#saving}
\CommentTok{#saveRDS(religionBaseline, }
\CommentTok{#file = "cosineAnalysis/models/religionBaseline.rds")}
\end{Highlighting}
\end{Shaded}

The only reason we need it is the evaluation of connection as a
predictor. Does including it in o the model improve the situation? To
investigate this, let's now build a model according to the following
specification:

\begin{align}
cosineDistance_i  & \sim dnorm(\mu_i, \sigma) \\
\mu_i & = m_{pw} + co_{con}\\
m_{pw} & ~ dnorm(1,.5) \\
co_{con} & ~ dnorm(0,1) \\
\sigma &\sim  dcauchy(0,1)
\end{align}

\noindent The idea now is that each connection type comes with its own
coefficient \(co\) that has impact on mean distances for protected words
taken separately.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(rethinking)}
\KeywordTok{options}\NormalTok{(}\DataTypeTok{buildtools.check =} \ControlFlowTok{function}\NormalTok{(action) }\OtherTok{TRUE}\NormalTok{ )}
\NormalTok{religionCoefs <-}\StringTok{ }\KeywordTok{ulam}\NormalTok{(}
  \KeywordTok{alist}\NormalTok{(}
\NormalTok{    cosineDistance }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(mu,sigma),}
\NormalTok{    mu <-}\StringTok{ }\NormalTok{m[pw] }\OperatorTok{+}\StringTok{ }\NormalTok{co[con],}
\NormalTok{    m[pw] }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    co[con] }\OperatorTok{~}\KeywordTok{dnorm}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    sigma }\OperatorTok{~}\StringTok{ }\KeywordTok{dcauchy}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)}
\NormalTok{  ),}
  \DataTypeTok{data =}\NormalTok{ religion,}
  \DataTypeTok{chains=}\DecValTok{2}\NormalTok{ , }\DataTypeTok{iter=}\DecValTok{8000}\NormalTok{ , }\DataTypeTok{warmup=}\DecValTok{1000}\NormalTok{, }
  \DataTypeTok{log_lik =} \OtherTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\normalsize

\noindent First, let's see if this model is really better in terms of
the Widely Acceptable Information Criterion (WAIC):

\vspace{1mm} \footnotesize

\begin{verbatim}
##                   WAIC SE dWAIC dSE pWAIC weight
## religionCoefs    -2328 93     0  NA    20      1
## religionBaseline -2283 95    45  17    16      0
\end{verbatim}

\normalsize

Clearly, it should be given weight 1 as compared to the baseline model.
So far, we've learned that the connection type actually has predictive
value. Let's take a look at the coefficient estimates:

\vspace{1mm} \footnotesize

\begin{verbatim}
##              mean        sd       5.5%      94.5%    n_eff    Rhat4
## co[1] -0.14956420 0.1151675 -0.3261650 0.03741930 294.2033 1.001449
## co[2] -0.09880543 0.1145024 -0.2736985 0.08813271 291.5044 1.001564
## co[3] -0.07282752 0.1133894 -0.2447778 0.11158986 287.7820 1.001627
## co[4] -0.03103179 0.1131420 -0.2034442 0.15268770 286.8283 1.001606
\end{verbatim}

\normalsize

\section{Dataset-level coefficients}\label{dataset-level-coefficients}

\noindent Let's plot them together with their highest posterior density
invervals, for the three topic groups.

\begin{center}
\begin{figure}[!htb]\centering
   \begin{minipage}{0.55\textwidth}
  \includegraphics[width=7cm]{../images/religionCoeffs.jpeg}
   \end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/religionGoogleCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/genderCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/genderGoogleCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/raceCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/raceGoogleCoeffs.jpeg}
   \end{minipage}
\end{figure}

\end{center}

\noindent What should strike us is that while the mean estimates of the
coefficients indeed do differ a bit, usually the highest posterior
density invervals all include zero, and so we do not have strong reasons
to say that, say, as far as the whole religion dataset is involved,
being associated indeed is connected with lower cosine distance. A
second striking observation is that the estimated impact for associated
stereotypes is quite often not too different from the estimated impact
of attributes associated with different stereotypes, and both are
sometimes not too far from the estimated impact for simply human
attributes. In general, once the uncertainty involved is taken seriously
by using control groups and statistical uncertainty estimation that does
not dispose of pointwise data, the picture which focuses only on
differences between means of means is too simplistic.

But this doesn't mean important differences for some protected words are
not there. For one thing, if you start with a word list that is very
uneven, the actually not so bad status of some of the protected words
might mask a pretty bad situation in which some other protected words
are. For comparison, let's see what a model focused on words related to
islam tells us.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#this is how we build the model}
\NormalTok{religion <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"cosineAnalysis/datasets/religionReddit.csv"}\NormalTok{)[}\OperatorTok{-}\DecValTok{1}\NormalTok{]}
\KeywordTok{colnames}\NormalTok{(religion) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"protectedWord"}\NormalTok{,}\StringTok{"wordToCompare"}\NormalTok{,}\StringTok{"wordClass"}\NormalTok{,}
                        \StringTok{"cosineDistance"}\NormalTok{,}\StringTok{"cosineSimilarity"}\NormalTok{,}\StringTok{"connection"}\NormalTok{)}
\KeywordTok{levels}\NormalTok{(religion}\OperatorTok{$}\NormalTok{wordClass) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"christian"}\NormalTok{,}\StringTok{"human"}\NormalTok{,}\StringTok{"jewish"}\NormalTok{,}\StringTok{"muslim"}\NormalTok{,}\StringTok{"neutral"}\NormalTok{)}
\NormalTok{muslimWords <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"imam"}\NormalTok{,}\StringTok{"islam"}\NormalTok{,}\StringTok{"mosque"}\NormalTok{,}\StringTok{"muslim"}\NormalTok{,}\StringTok{"quran"}\NormalTok{)}
\NormalTok{muslim <-}\StringTok{ }\NormalTok{religion }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(protectedWord }\OperatorTok{%in%}\StringTok{ }\NormalTok{muslimWords)}
\NormalTok{muslim}\OperatorTok{$}\NormalTok{protectedWord <-}\StringTok{ }\KeywordTok{droplevels}\NormalTok{(muslim}\OperatorTok{$}\NormalTok{protectedWord)}
\NormalTok{muslim}\OperatorTok{$}\NormalTok{pw <-}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(muslim}\OperatorTok{$}\NormalTok{protectedWord)}
\NormalTok{muslim}\OperatorTok{$}\NormalTok{con <-}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(muslim}\OperatorTok{$}\NormalTok{connection)}
\NormalTok{muslim}\OperatorTok{$}\NormalTok{pwFactor <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(muslim}\OperatorTok{$}\NormalTok{protectedWord, muslim}\OperatorTok{$}\NormalTok{connection))}
\NormalTok{muslim}\OperatorTok{$}\NormalTok{pwIndex <-}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(muslim}\OperatorTok{$}\NormalTok{pwFactor)}

\NormalTok{islamCoefs <-}\StringTok{ }\KeywordTok{ulam}\NormalTok{(}
  \KeywordTok{alist}\NormalTok{(}
\NormalTok{    cosineDistance }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(mu,sigma),}
\NormalTok{    mu <-}\StringTok{ }\NormalTok{m[pw] }\OperatorTok{+}\StringTok{ }\NormalTok{co[con],}
\NormalTok{    m[pw] }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    co[con] }\OperatorTok{~}\KeywordTok{dnorm}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    sigma }\OperatorTok{~}\StringTok{ }\KeywordTok{dcauchy}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)}
\NormalTok{  ),}
  \DataTypeTok{data =}\NormalTok{ muslim,}
  \DataTypeTok{chains=}\DecValTok{2}\NormalTok{ , }\DataTypeTok{iter=}\DecValTok{10000}\NormalTok{ , }\DataTypeTok{warmup=}\DecValTok{1000}\NormalTok{, }\DataTypeTok{cores =} \DecValTok{4}\NormalTok{,}
  \DataTypeTok{log_lik =} \OtherTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\normalsize

Let's take a look at the coefficients:

\vspace{1mm} \footnotesize

\begin{verbatim}
##                mean        sd       5.5%      94.5%    n_eff    Rhat4
## co[1] -0.1979930035 0.1708785 -0.4682696 0.07634977 1789.894 1.003445
## co[2] -0.0334215769 0.1687587 -0.3021954 0.23720938 1738.720 1.003575
## co[3] -0.0192492753 0.1675754 -0.2840596 0.24860974 1732.907 1.003755
## co[4] -0.0003911363 0.1670610 -0.2661047 0.26815172 1723.758 1.003837
\end{verbatim}

\normalsize

\begin{center}
\includegraphics[width=8cm]{../images/islamCoeffs.jpeg}
\end{center}

\normalsize

While muslim words were unusual in the sense that the disparity between
associated attributes and others is stronger, the evidence is still not
conclusive. This is because the variation even within islam-related
words is large enough (and sample sizes sufficiently small) for all the
highest posterior density invervals to still include zeros.

So, it seems, taking a closer look does seem to make a difference. The
question is, what happens if we do take a close look at the level of
protected words?

\chapter{Protected-word level
analysis}\label{protected-word-level-analysis}

\section{Model structure and
assumptions}\label{model-structure-and-assumptions}

Let's turn then to data analysis that takes a look at protected words
separately. This time for each combination of a protected word and a
connection status we will have a separate mean cosine distance estimate,
each coming with its own highest posterior density interval. This means
we will use indices that are result from all such combinations (and then
we will split them up in the model precis to build visualisation, feel
free to look at the \texttt{visualiseStats.R} script for details).

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{options}\NormalTok{(}\DataTypeTok{buildtools.check =} \ControlFlowTok{function}\NormalTok{(action) }\OtherTok{TRUE}\NormalTok{ ) }\CommentTok{#removes install pop-up request}
\NormalTok{religion}\OperatorTok{$}\NormalTok{pwFactor <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(religion}\OperatorTok{$}\NormalTok{protectedWord, }\StringTok{"-"}\NormalTok{, religion}\OperatorTok{$}\NormalTok{connection))}
\NormalTok{religion}\OperatorTok{$}\NormalTok{pwIndex <-}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(religion}\OperatorTok{$}\NormalTok{pwFactor)}

\NormalTok{religionSeparate <-}\StringTok{ }\KeywordTok{ulam}\NormalTok{(}
  \KeywordTok{alist}\NormalTok{(}
\NormalTok{    cosineDistance }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(mu,sigma),}
\NormalTok{    mu <-}\StringTok{ }\NormalTok{c[pwIndex],}
\NormalTok{    c[pwIndex] }\OperatorTok{~}\StringTok{ }\KeywordTok{dnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    sigma }\OperatorTok{~}\StringTok{ }\KeywordTok{dcauchy}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)}
\NormalTok{  ),}
  \DataTypeTok{data =}\NormalTok{ religion,}
  \DataTypeTok{chains=}\DecValTok{2}\NormalTok{ , }\DataTypeTok{iter=}\DecValTok{10000}\NormalTok{ , }\DataTypeTok{warmup=}\DecValTok{1000}\NormalTok{,}
  \DataTypeTok{start=}\KeywordTok{list}\NormalTok{(}\DataTypeTok{no =} \DecValTok{1}\NormalTok{, }\DataTypeTok{a =} \DecValTok{0}\NormalTok{, }\DataTypeTok{d =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sigma=} \FloatTok{.3}\NormalTok{), }\DataTypeTok{log_lik =}  \OtherTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\normalsize

\noindent Let's see if the individualized model does better than the
previous models in light of WAIC which does add penalty for the number
of parameters.

\vspace{1mm} \footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{compareBaselineCoefsSeparate<-}\StringTok{ }\KeywordTok{readRDS}\NormalTok{(}\StringTok{"../datasets/compareBaselineCoefsSeparate.rds"}\NormalTok{)}
\NormalTok{compareBaselineCoefsSeparate}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                   WAIC SE dWAIC dSE pWAIC weight
## religionSeparate -2400 93     0  NA    60      1
## religionCoefs    -2328 93    72  29    20      0
## religionBaseline -2283 95   117  37    16      0
\end{verbatim}

\normalsize

\section{Protected classes in Reddit and Google
embeddings}\label{protected-classes-in-reddit-and-google-embeddings}

It seems that we do want to prefer this model, despite its relative
complication. Now, what does it tell us about the protected words? Let's
visualise the predicted means together with 89\% highest posterior
density intervals.

\includegraphics[width=14cm]{../images/visReligionReddit.png}

Before we move on, let's perform analogous analyses for the remaining
types of supposed bias: gender and race (the model building is
analogous).

\includegraphics[width=14cm]{../images/visGenderReddit.png}

\includegraphics[width=14cm]{../images/visRaceReddit.png}

We first encountered the GoogleNews word embedding model in Bolukbasi et
al. (\protect\hyperlink{ref-Bolukbasi2016Man}{2016}). They use it for
their calculations as written in
\href{https://github.com/tolga-b/debiaswe}{Tolga Code}. We decided to
compare the results obtained with the use of Reddit L2 model and the
ones that we got by GoogleNews model. The details of the model can be
found here \href{https://code.google.com/archive/p/word2vec/}{Google
News model}. One of the main differences between the models is that
Reddit word embeggings have 50 dimensions and GoogleNews word embeddings
have 300 dimensions. As the dimension is increasing, the vector can
capture much more information but this is not always the case. According
to some researchers the amount of dimensions should be chosen by taking
into account the type of corpus and its features. In our case both
Reddit and GoogleNews models are already trained and ready to use so we
do not analyse further the choice of their hyperparameters.

\includegraphics[width=14cm]{../images/visReligionGoogle.png}

\includegraphics[width=14cm]{../images/visGenderGoogle.png}

\includegraphics[width=14cm]{../images/visRaceGoogle.png}

\chapter{The role of debiasing}\label{the-role-of-debiasing}

We also applied to the word embeddings hard debiasing method found in
Manzini et al.
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}). As it was
mentioned before, debiasing consists of two components: identifying the
bias subspace and then removing this vector subspace from the chosen
embeddings. The aim is to increase the cosine distance between protected
words and the set of attributes. After debiasing the Reddit word
embeddings we calculated the cosine distance again to see if there are
any significant changes in terms of the similarity between protected
words and the attributes.

\section{Dataset-level coefficients after
debiasing}\label{dataset-level-coefficients-after-debiasing}

First, let's look at coefficient estimated for the whole datasets, as
compared to their estimation prior to debiasing:

\begin{center}
\begin{figure}[!htb]\centering
   \begin{minipage}{0.55\textwidth}
  \includegraphics[width=7cm]{../images/religionCoeffs.jpeg}
   \end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/debiasedReligionRedditCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/genderCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/debiasedGenderRedditCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/raceCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/debiasedRaceRedditCoeffs.jpeg}
   \end{minipage}
\end{figure}

\end{center}

One may observe very slight changes in the estimated coeﬀicients. It
seems that in religion dataset the changes are the most significant as
the mean moves towards zero which stands for no similarity between
words.

\section{Protected classes after
debiasing}\label{protected-classes-after-debiasing}

Now, perhaps, the effects of debiasing will be better appreciated if we
look at the level of protected words. After all, the hope is, the
situation of extremely ill-positioned protected words have improved?

\includegraphics[width=14cm]{../images/visDebReligionReddit.png}

\includegraphics[width=14cm]{../images/visDebGenderReddit.png}

\includegraphics[width=14cm]{../images/visDebRaceReddit.png}

One may notice that in a religion data set most of the distances for
\texttt{associated} words slightly increased. Some of the distances from
\texttt{different} class are still smaller than for the
\texttt{associated} class. The distances for all classes seem in most
cases to cluster together even after debiasing. The reason may be that
uncertainty is quite high in \texttt{associated} and \texttt{different}
class because the number of samples is so small. Additionally in some
cases distances from \texttt{human} class are smaller than for the
\texttt{assosiated} class which also makes it less clear how to
interpret the results, if the distances are really able to catch the
bias presence.

It is worth paying attention to the gender debiased dataset. The change
is really minor, still zero out of HPDI range. The \texttt{different}
attributes still have high similarity value. It seems that the distances
for gender protected words are still clustered together which suggests
that this method is not able to catch the complex bias presence. One
could conclude that co-occurrence of certain terms does not always mean
direct connection in terms of terms associations. The fact that some of
stereotypically associated with female attributes still have high
similarity with male protected words suggest that other techniques
should be applied to detect and remove the bias.

At the same time, there is a small improvement in race dataset. In some
cases the distances just cluster together. In general distances for
\texttt{assosiated} class are still higher than those for the rest of
classes. One may assume that there reason for so small change is that
either the metric does not indicate properly biases or that the issue is
more complex and subtracting subspaces is not enough.

In Manzini et al.
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}) it is
mentioned that debiasing with this method that is based on bias
component removal may be insufficient. The reason is that words
containing similar biases may still cluster together even after
debiasing.

\chapter{Discussion and summary}\label{discussion-and-summary}

We propose the use of Bayesian methods to measure uncertainty in bias
detection. There are a few advantages of this method. One of them is
that including uncertainty enables one to directly observe the influence
of sample sizes. Analyzing individual words and connection coefficients,
one may notice how \texttt{neutral} words have smaller uncertainty
intervals and \texttt{different} or \texttt{associated} quite the
opposite. One of the reasons for such outcome is that we used
approximately 230 neutral words and only between 11-25 (the number
varies from class to class) stereotypical attributes from Manzini et al.
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}) article. In
our approach, we also pay attention to the distribution and details
regarding anomalous values. With the use of simple visualizations that
we introduced before, we were able to indicate suspicious cosine
distance values. Additionally, we compare in details how the cosine
distance values and uncertainty change after the debiasing. One can
verify then how the individual vectors changed and if it is what was
expected. Our analysis with the use of Bayesian method gave us new ideas
and hypothesis concerning not only the bias detection method but also
the efficiency of debiasing itself.

We created a summary table for each of the datasets: Reddit, Reddit
Debiased, and Google word embeddings. Let us first analyse the general
observations from estimated coefficients mean introduced in 3.6.
DATASET-LEVEL COEFFICIENTS. For Google embeddings the HPDI for all
classes coefficients (associated, different, human, and none) has an
interval that includes zero. This can lead one to a conclusion that the
impact for associated, different, human and neutral attribute is, when
averaged, quite similar. This indicates how including the uncertainty
may change the use and interpretation of Manzini et al.
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}) MAC metric.
It seems that if one focuses only on differences between means of means,
it is too simplistic. In case of Reddit word embeddings the situation is
similar although HPDI interval is below 0 for Gender class when looking
at \texttt{associated} and \texttt{different} mean coefficient. This can
suggest that there is indeed slightly stronger impact of these
attributes. One should also notice how in general \texttt{associated}
and \texttt{different} coefficients have quite similar HPDI range, the
highest observed absolute difference is equal to only 0.1. This suggests
that again the impact of associated attributes and difference ones is
not clear at first sight when looking at averaged coefficients. Finally,
let's compare the HPDI intervals for Reddit and Reddit debiased
datasets. For Religion and Race dataset there is a minor shift (in
absolute values the highest change is equal to approximately 0.1) of the
mean coefficients towards zero. However for Gender dataset there is no
significant change. This is of course the general look at the data,
let's now analyze the individual words.

In Reddit table one can observe that for Religion class the cosine
distance results for the associated attributes are for approximately
60\% of the words close to human attributes as well. This can suggest
that in some cases words concerning humans can have higher similarity
with some protected words independently if they are stereotypical or
neutral-human words. One should also pay attention to the fact that for
all of the \texttt{associated} and \texttt{different} attributes in
Religion class the uncertainty interval overlaps at some point. What is
even more surprising is that for protected words, such as ``torah'' the
\texttt{associated} attribute has the the cosine distance slightly over
1, which means no positive similarity! If the protected words that we
chose do not have high similarity with harmful associated attributes,
then one should consider at least three scenarios. The first one is that
the choice of the protected words and attributes may be corrupted. The
second one is that the metric is not able to catch the hidden bias
properly. The third one is that there is actually no bias between the
words. Regardless of which scenario one considers, it is essential to
take a look at the individual values before averaging them or
aggregating in other ways. It seems that using Bayesian method can
enhance the process of verifying the hypotheses concerning the choice of
protected words and attributes.

Surprisingly in Gender class one can observe high cosine similarity
values between some female stereotypical professions and male protected
words. If a word stereotypically associated with females has low cosine
distance to male protected words, then one should investigate the issue
further. The reasons for this unexpected cosine distance may lie in the
frequency of appearance of the protected words in the raw data. Some of
the groups (like Muslim people or females) may have less representation
in the data that is taken as an input for word embeddings. Therefore, if
we assume that the MAC detects actually co-occurrence only, it makes
sense that they can have high similarity with associated attributes and
lower similarity with different ones. At the same time male protected
words and other religions can have high similarity with different
attributes because they have greater representation in the dataset in
general and occur close to much more concepts. Cosine distance seems to
capture the information on the co-occurrence of words and not on the
semantic similarity strictly speaking.

In Google table one may notice interesting observations as well.
Although GoogleNews dataset is larger, trained on different data sources
and with the use of more dimensions, some of the results are similar to
the ones obtained for the Reddit corpus. For Religion case almost all of
the values for \texttt{associated} class have an intersection with
\texttt{different} class as well. In the case of Race it is 100\% of the
available words. This indicates again that it is not clear how the
metric should be used. Quite different situation takes place for Gender
data where the similarity between \texttt{associated} and
\texttt{different} is present mostly for the male protected words. This
means that male words have high similarity with both male stereotypical
professions and female ones. However, for females the similarity is high
mostly only for the female stereotypical professions. This observation
could not be make when using MAC metric only as it requires
investigating the individual words.

When analyzing the Reddit debiased results one should ask a question of
the change of the cosine distance values. It seems that not all of the
protected words are treated equally when performing debiasing. In
religion data the values for \texttt{associated} class moved towards 0
for most of the words except the ones for Islam religion where still the
\texttt{associated} class has quite lower cosine distance than the
\texttt{different} one. Similar situation takes place in Gender data
where female words have still much lower cosine distance values for
\texttt{associated} class than for the \texttt{different} one. In the
case of males it is mostly almost the same interval for
\texttt{associated} and \texttt{different}. As the Gender data is quite
specific as the attributes are not harmful adjectives but (in an ideal
world) neutral professions, the aim (if we follow MAC assumptions)
should actually be to make the cosine distance same for both female and
male protected words. However, as we pointed out the situation after
debiasing can sometimes be better only for one protected group which is
not the desired outcome. In the case of Religion data even after
debiasing, all of the protected words have intersections between
\texttt{associated} and \texttt{different} class and similarity greater
than 0. As all of the attributes in religion data are negative and
harmful stereotypes, one should not aim at making the distance between
protected words and \texttt{associated}, and \texttt{different} class
the same but rather to move it towards 0.

Let's summarize the results of the bias detection methods analysis. As
it was presented above, using the bias detection methods without
involving the uncertainty may bring the risk of limited insight into the
data. One should be concerned with the limitations of the metric that
evaluates the debiasing by the mean of the averaged mean approach.
Additionally, one cannot be sure if the bias is still preserved after
debiasing. The fact that all of the cosine distances for protected words
and harmful attributes moved slightly to the right, does not mean that
the bias is removed. It is shown in articles such as Gonen \& Goldberg
(\protect\hyperlink{ref-Gonen2019Lipstick}{2019}), that the bias can
hide in the vector geometry and preserve even after applying popular
debiasing methods. One of the reasons for that may be the fact that
relative differences between words may be preserved even after
debiasing. Another thing is that the choice of word lists used to verify
the metric effectiveness is not well justified. The sample size is very
small which leads to large HPDI. Some of the protected words and
attributes originate from old articles and it seems that the bias
phenomenon is quite dynamic and the data should be as up-to-date as it
is possible. One could for example research the IAT examples to verify
the effectiveness of MAC metric. One should remember that there is no
clear interpretation for the values obtained with MAC metric from
Manzini et al.
(\protect\hyperlink{ref-Manzini2019blackToCriminal}{2019}). One may
assume that if the cosine distance is close to 1 then it is a desired
outcome as it means, according to cosine distance assumptions, that
there is almost no similarity between the words. However what does it
mean to be close to 0? If the averaged cosine distance is equal to 0.8,
then should we still debias it? It is unclear what the criteria are. On
one hand, it seems to be beneficial when the outcome is simplified as it
is easier to compare results with one value per set. On the other hand,
it is prone to misunderstanding of how to interpret the results and what
threshold to assume.

The bottom line is that if we want to take bias seriously, so should we
approach the uncertainty involved in our estimations. There is no
replacement for proper statistical evaluation that does not discard
information about the uncertainty involved, larger word lists are
needed, and visualization of the results for particular protected
classes provides much better guidance than chasing a single metric based
on a means of means.

Bibliography:

\hypertarget{refs}{}
\hypertarget{ref-Bolukbasi2016Man}{}
Bolukbasi, T., Chang, K., Zou, J. Y., Saligrama, V., \& Kalai, A.
(2016). Man is to computer programmer as woman is to homemaker?
Debiasing word embeddings. \emph{CoRR}, \emph{abs/1607.06520}. Retrieved
from \url{http://arxiv.org/abs/1607.06520}

\hypertarget{ref-Gonen2019Lipstick}{}
Gonen, H., \& Goldberg, Y. (2019). Lipstick on a pig: Debiasing methods
cover up systematic gender biases in word embeddings but do not remove
them. \emph{CoRR}, \emph{abs/1903.03862}. Retrieved from
\url{http://arxiv.org/abs/1903.03862}

\hypertarget{ref-Caliskan2017Semantics}{}
Islam, A. C., Bryson, J. J., \& Narayanan, A. (2016). Semantics derived
automatically from language corpora necessarily contain human biases.
\emph{CoRR}, \emph{abs/1608.07187}. Retrieved from
\url{http://arxiv.org/abs/1608.07187}

\hypertarget{ref-Manzini2019blackToCriminal}{}
Manzini, T., Lim, Y. C., Tsvetkov, Y., \& Black, A. W. (2019). Black is
to criminal as caucasian is to police: Detecting and removing multiclass
bias in word embeddings.

\hypertarget{ref-Mehrabi2019Survey}{}
Mehrabi, N., Morstatter, F., Saxena, N., Lerman, K., \& Galstyan, A.
(2019). A survey on bias and fairness in machine learning. \emph{CoRR},
\emph{abs/1908.09635}. Retrieved from
\url{http://arxiv.org/abs/1908.09635}

\hypertarget{ref-Nissim2019Fair}{}
Nissim, M., Noord, R. van, \& Goot, R. van der. (2019). Fair is better
than sensational: Man is to doctor as woman is to doctor. \emph{CoRR},
\emph{abs/1905.09866}. Retrieved from
\url{http://arxiv.org/abs/1905.09866}

\hypertarget{ref-Venkat2018Curse}{}
Venkat, N. (2018). The curse of dimensionality: Inside out.

\end{document}