### 2021-07-21 paperDraft/paperDraft.tex
Adds small changes

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  10pt,
  dvipsnames,enabledeprecatedfontcommands]{scrartcl}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Taking uncertainty in word embedding bias estimation seriously: a bayesian approach},
  pdfauthor={Alicja Dobrzeniecka and Rafal Urbaniak},
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%\documentclass{article}

% %packages
 \usepackage{booktabs}

\usepackage{multirow}

\usepackage{graphicx}
\usepackage{longtable}
\usepackage{ragged2e}
\usepackage{etex}
%\usepackage{yfonts}
\usepackage{marvosym}
\usepackage[notextcomp]{kpfonts}
\usepackage{nicefrac}
\newcommand*{\QED}{\hfill \footnotesize {\sc Q.e.d.}}
\usepackage{floatrow}

\usepackage[textsize=footnotesize]{todonotes}
%\linespread{1.5}

\setlength{\parindent}{10pt}
\setlength{\parskip}{1pt}

%language
\usepackage{times}
\usepackage{t1enc}
%\usepackage[utf8x]{inputenc}
%\usepackage[polish]{babel}
%\usepackage{polski}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx}
\usepackage[scaled=0.88]{helvet}

%AMS
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}

\usepackage{geometry}
 \geometry{a4paper,left=35mm,top=20mm,}

%environments
\newtheorem{fact}{Fact}

%abbreviations
\newcommand{\ra}{\rangle}
\newcommand{\la}{\langle}
\newcommand{\n}{\neg}
\newcommand{\et}{\wedge}
\newcommand{\jt}{\rightarrow}
\newcommand{\ko}[1]{\forall  #1\,}
\newcommand{\ro}{\leftrightarrow}
\newcommand{\exi}[1]{\exists\, {_{#1}}}
\newcommand{\pr}[1]{\mathsf{P}(#1)}
\newcommand{\cost}{\mathsf{cost}}

\newcommand{\odds}{\mathsf{Odds}}
\newcommand{\ind}{\mathsf{Ind}}
\newcommand{\nf}[2]{\nicefrac{#1\,}{#2}}
\newcommand{\R}[1]{\texttt{#1}}
\newcommand{\prr}[1]{\mbox{$\mathtt{P}_{prior}(#1)$}}
\newcommand{\prp}[1]{\mbox{$\mathtt{P}_{posterior}(#1)$}}

\newtheorem{q}{\color{blue}Question}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

%technical intermezzo
%---------------------

\newcommand{\intermezzoa}{
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Starts}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}\nopagebreak 
	}

\newcommand{\intermezzob}{\nopagebreak 
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Ends}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}
	}
%--------------------

\newtheorem*{reply*}{Reply}
\usepackage{enumitem}
\newcommand{\question}[1]{\begin{enumerate}[resume,leftmargin=0cm,labelsep=0cm,align=left]
\item #1
\end{enumerate}}

\usepackage{float}

% \setbeamertemplate{blocks}[rounded][shadow=true]
% \setbeamertemplate{itemize items}[ball]
% \AtBeginPart{}
% \AtBeginSection{}
% \AtBeginSubsection{}
% \AtBeginSubsubsection{}
% \setlength{\emergencystretch}{0em}
% \setlength{\parskip}{0pt}

\usepackage[authoryear]{natbib}

%\bibliographystyle{apalike}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Taking uncertainty in word embedding bias estimation seriously: a
bayesian approach}
\author{Alicja Dobrzeniecka and Rafal Urbaniak}
\date{}

\begin{document}
\maketitle

\tableofcontents

\hypertarget{cosine-based-measures-of-bias}{%
\section{Cosine-based measures of
bias}\label{cosine-based-measures-of-bias}}

Modern Natural Language Processing (NLP) models are used to complete
various tasks such as providing email filters, smart assistants, search
results, language translations, text analytics and so on. All of them
need as an input words represented by means of numbers which is
accomplished with word embeddings. It seems that in the learning process
these models can learn implicit biases that reflect harmful
stereotypical thinking. One of the sources of bias in NLP can be located
in the way the word embeddings are made. There is a considerable amount
of literature available on the topic of bias detection and mitigation in
NLP models.

One of the first measures in the discussion has been developed by
(\textbf{bolukbasi2016man?}). There, the gender bias of a word \(w\) is
understood as its projection on the gender direction
\(\vec{w} \cdot (\overrightarrow{he} - \overrightarrow{she})\) (the
gender direction is the top principal compontent of ten gender pair
difference vectors). The underlying idea is that no bias is present if
non-explicitly gendered words are in equal distance to both elements in
all explicitly gender pairs. Given the (ideally) gender neutral words
\(N\) and the gender direction \(g\) the direct gender bias is defined
as the average distance of the words in \(N\) from \(g\) (\(c\) is a
parameter determining how strict we want to be): \begin{align}
\mathsf{directBias_c(N,g)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},g)\vert^c}{\vert N \vert }
\end{align}

The use of projections has been ciriticized for instance by
(\textbf{Gonen2019lipstick?}), who point out that while gender-direction
might be an indicator of bias, it is only one possible manifestation of
it, and reducing a projection of words might be insufficient. For
instance, ``math'' and ``delicate'' might be in equal distance to both
explicitly gendered words while being closer to quite different
stereotypical attribute words. Further, the authors point out that most
word pairs preserve similarity under debiasing meant to minimize
projection-based bias.\footnote{(\textbf{bolukbasi2016man?}) use also
  another method which involves analogies and their evaluations by human
  users on Mechanical Turk. It is discussed and criticized in
  (\textbf{Nissim2020fair?}).}

To measure bias in word embeddings,
(\textbf{Caliskan2017semanticsBiases?}) proposed the Word Embedding
Association Test (WEAT). The idea is that the measure of biases between
two sets of target words, \(X\) and \(Y\), (we call them protected
words) should be quantified in terms of the cosine similarity between
the protected words and attribute words coming from two sets of
stereotype attribute words, \(A\) and \(B\) (we'll call them
attributes). For instance, \(X\) might be a set of male names, \(Y\) a
set of female names, \(A\) might contain stereotypically male-related
career words, and \(B\) stereotypically female-related family words.
WEAT is a modification of the Implicit Association Test (IAT)
(\textbf{Nosek2002harvesting?}) used in psychology and uses almost the
same word sets, allowing for a \emph{prima facie} sensible comparison
with bias in humans. If the person's attitude towards given pair of
concept is to be interpreted as neutral, there should be no noticeable
task completion time difference, and the final value from the formula
should be around 0. Let \(f\) be a similarity measure (usually, cosine
similarity). The association difference for a term \(t\) is:
\begin{align}
s(t,A,B) & = \frac{\sum_{a\in A}f(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}f(t,b)}{\vert B\vert}
\end{align} \noindent then, the association difference between \(A\) a
\(B\) is: \begin{align}
s(X,Y,A,B) & = \sum_{x\in X} s(x,A,B) -  \sum_{y\in Y} s(y,A,B)
\end{align} \noindent \(s(X,Y,A,B)\) is the statistic used in the
signifcance test, and the \(p\)-value obtained by bootstrapping: it is
the frequency of \(s(X_i,Y_i,A,B)>s(X,Y,A,B)\) for all equally sized
partitions \(X_i, Y_i\) of \(X\cup Y\). The effect size is computed by
normalizing the difference in means as follows: \begin{align}
bias(A,B) & = \frac{
\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
}
\end{align}

(\textbf{Caliskan2017semanticsBiases?}) show that significant
biases---thus measured--- similar to the ones discovered by IAT can be
discovered in word embeddings. (\textbf{Lauscher2019multidimensional?})
extended the methodology to a multilingual and cross-lingual setting,
arguing that using Euclidean distance instead of similarity does not
make much difference, while the bias effects vary greatly across
embedding models (interestingly, with social media-text trained
embeddings being less biased than those based on Wikipedia).

A similar methodology is employed by (\textbf{Garg2018years?}), who
employ word embeddings trained on corpora from different decades to
study the shifts in various biases. For instance, to compute the
occupational embeddings bias for women the authors first compute the
average vector of vector emeddings of words that represent women
(e.g.~``she,'' ``female''), then calculate the Euclidean distance
between this mean vector and words for occupations. Then they take the
mean of these distances and subtract from it the analogously obtained
mean for the average vector of vector embeddings of words that represent
men. Formally they take the relative norm distance between \(X\) and
\(Y\) to be: \begin{align}
\textsf{relative norm distance} & = \sum_{v_m\in M} \vert \vert v_m - v_X\vert \vert_2 - \vert v_m - v_Y\vert \vert_2
\end{align} \noindent where the norm used is Euclidean, and \(v_X\) and
\(x_Y\) are average vectors for sets \(X\) and \(Y\) respectively.

Manzini, Lim, Tsvetkov, \& Black (2019) modify WEAT to a multi-class
setting, introducing Mean Average Cosine similarity as a measure of bias
(in fact, in the paper they report distances rather than similarities).
Let \(T = \{t_1, \dots, t_k\}\) be a class of protected word embeddings,
and let each \(A_j\in A\) be a set of attributes stereotypically
associated with a protected word). Then: \begin{align}
S(t_i, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) \\
MAC(T,A) & = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j)
\end{align} That is, for each protected word \$T and each attribute
class, they first take the mean for this protected word and all
attributes in a given attribute class, and then take the mean of thus
obtained means for all the protected words. The t-tests they employ are
run on average cosines used to calculate MAC.

\hypertarget{some-methodological-problems}{%
\section{Some methodological
problems}\label{some-methodological-problems}}

Such approaches, however, are not unproblematic. One may cosinder the
selection of attributes for the bias research. For one thing, the word
lists are often fairly small, and the papers do not discuss the impact
of word list sizes on the uncertainty involved. We will argue that the
word list sizes are too small to yield strong conclusions, given the
variances involved.

Crucially, these approaches use means of mean average cosine
similarities to measure similarity between protected word and harmful
stereotypes. If one takes a closer look at the individual values that
are taken for the calculations, it turns out that there are quite a few
outliers and surprisingly dissimilar words. This issue will become
transparent when we inspect the visualizations of individual cosine
distances, following the idea that one of the first step to understand
data is to look at it.

With such a method the uncertainty involved is not really considered
which makes it even more difficult to give reasonable interpretations of
the results. We propose the use of Bayesian method to obtain some
understanding of the influence the uncertainty has on the interpretation
of final results.

We also observed that in the paper there is no word class distinction
and no control group. In the original paper, words from all three
religions were compared against all of the stereotypes, which means that
there was no distinction between cases in which the stereotype is
associated with a given religion, as opposed to the situation in which
it is associated with another one. Not all of the stereotypical words
have to be considered as harmful for all of the religions. One should
investigate the religions separately as some of them may have stronger
harmful associations that others. One should also include control groups
to have a way of comparing the stereotypical results with neutral or
human-like words. Later in the text we will explain in details reasons
for introducing control groups. In our analysis, we distinguish between
stereotypes associated with a given group, stereotypes associated with
different groups, and control groups: neutral words and stereotypes-free
human predicates.

The interpretation of the results is also a challenge. Assuming for a
moment that the value of multi-class cosine distance is correct, one may
question the interpretation. Manzini, Lim, Tsvetkov, \& Black (2019)
summarize the averages of cosine distance per group (gender, race,
religion). For now let us focus now on analyzing the values relating to
religious biases. Here is the relevant fragment of table:

\begin{longtable}[]{@{}ll@{}}
\toprule
Religion Debiasing & MAC \\
\midrule
\endhead
Biased & 0.859 \\
Hard Debiased & 0.934 \\
Soft Debiased (\(\lambda\) = 0.2) & 0.894 \\
\bottomrule
\end{longtable}

MAC stand for mean average cosine similarity, although in reality the
the table contains mean distances. What may attract attention is the
fact that the value of cosine distance in ``Biased'' category is already
quite high even before debiasing. High cosine distance indicates low
cosine similarity between values. One could think that average cosine
similarity equal to approximately 0.141 is not significant enough to
consider it as bias. However the authors aim to mitigate ``biases'' in
vectors with such great distance to make it even larger.
Methodologically the question is, on what basis is this small similarity
still considered as a proof of the presence of bias, and whether these
small changes are meaningful. This is in general the problem of scale
and the lack of universal intervals. In contrast, statistical intervals
will help us decide whether a given cosine similarity is high enough to
consider the words to be more similar than if we chose them at random.
We will use highest posterior density intervals, in line with Bayesian
methodology.

One may consider the issue of the curse of dimensionality. In our case,
the curse of dimensionality may take place when there is an increase in
the volume of data that results in adding extra dimensions to the
Euclidean space. As the number of features increases, it may be harder
and harder to obtain useful information from the data using the
available algorithms. Using cosine similarity in high dimensions in word
embeddings may be prone to the curse of dimensionality. According to
Venkat (2018) there are reasons to consider this phenomenon when
searching for word similarities in higher dimensions. An experiment is
conducted that aims at showing how the similarity values and variation
change as the number of dimensions increases. The hypothesis made in the
paper states that two things will happen as the number of dimensions
increase. First, the effort required to measure cosine similarity will
be greater, and two, the similarity between data will blur out and have
less variation. The authors generate random points with increasing
number of dimensions where each dimension of a data point is given a
value between 0 and 1. Then they pick one vector at random from each
dimension class and calculate the cosine similarity between the chosen
vector and the rest of the data. Then they check how the variation of
values changes as the number of dimensions increases. It seems like the
more dimensions there are, the smaller the variance and therefore it is
less obvious how to interpret the resulting cosine similarities. Maybe
the scale should be adjusted to the number of dimensions and variance so
that it still gives us sensible information about data.

\hypertarget{the-problem-with-means-of-means}{%
\section{The problem with means of
means}\label{the-problem-with-means-of-means}}

The measures described all calculate means of means and their authors
run statistical tests on sets of means. This, however, is problematic
for two related reasons. One, by pre-averaging data we throw away
information about sample sizes. For the former point, think about
proportions: 10 out of 20 and 2 out of 4 give the same mean, but you
would obtain more information by making the former observation rather
than by making the latter. Two, when we pre-average, we remove
variation, and so pre-averaging tend to manufacture false confidence. We
will have more to say about whether this happens in the case of
applications of MAC, for now let's go over a simple example to make the
conceptual point clear.

To illustrate let's employ the formulas used by
(\textbf{Caliskan2017semanticsBiases?}) in a simple example.
Conceptually, all such tests come up with rather short lists of
protected words and rather short lists of stereotypical attributes.
Clearly, these are not complete list. So let's treat them as samples
from richer pools of stereotypical predicates and let's take the
uncertainty and variation involved seriously.

Consider a simple situation in which there are two protected classes,
\(X=\{t_1,t_2\}\) and \(Y=\{t_3,t_4\}\) and two five-element attribute
sets \(A\) and \(B\).

First, we play around with a scenario in which all the protected terms
are on average equally unsimilar to both sets (\(\mu =0\)) with standard
deviation of \(.05\). Let's randomly draw similarity scores and plot the
results with group means plotted as vertical lines.

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\NormalTok{t1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t3 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t4 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-2-1} \end{center}

\noindent When you look at the datapoints, do you have the impression of
a strong bias here? We wouldn't think so. But now let's run the
calculations from (\textbf{Caliskan2017semanticsBiases?}).

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{s }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (table)\{ }\FunctionTok{mean}\NormalTok{(table}\SpecialCharTok{$}\NormalTok{A) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(table}\SpecialCharTok{$}\NormalTok{B)\}}
\NormalTok{factor }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1),}\FunctionTok{s}\NormalTok{(t2),}\FunctionTok{s}\NormalTok{(t3),}\FunctionTok{s}\NormalTok{(t4)))}
\NormalTok{numerator }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1),}\FunctionTok{s}\NormalTok{(t2)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3),}\FunctionTok{s}\NormalTok{(t4))}
\FunctionTok{print}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{factor =}\NormalTok{ factor,}\AttributeTok{numerator =}\NormalTok{ numerator, }\AttributeTok{bias =}\NormalTok{ numerator }\SpecialCharTok{/}\NormalTok{ factor))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $factor
## [1] 0.02342637
## 
## $numerator
## [1] 0.04275325
## 
## $bias
## [1] 1.825005
\end{verbatim}

\normalsize

\noindent This should make us pause. We know these were points randomly
drawn from distributions where there is no difference in means. Yet, the
calculated effect size is 1.82, whereas the largest effect size reported
by (\textbf{Caliskan2017semanticsBiases?}) is 1.81!

Interestingly, if we repeat the drawing 10000 times, each time
calculating the bias, it turns out that with this variance and sample
size, pretty much anything can happen.

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{biasesNull }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\DecValTok{10000}\NormalTok{)\{}
\NormalTok{t1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t3 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t4 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}

\NormalTok{factor }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1),}\FunctionTok{s}\NormalTok{(t2),}\FunctionTok{s}\NormalTok{(t3),}\FunctionTok{s}\NormalTok{(t4)))}
\NormalTok{numerator }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1),}\FunctionTok{s}\NormalTok{(t2)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3),}\FunctionTok{s}\NormalTok{(t4))}
\NormalTok{biasesNull[i]  }\OtherTok{\textless{}{-}}\NormalTok{ numerator }\SpecialCharTok{/}\NormalTok{ factor}
\NormalTok{\}}
\FunctionTok{ggplot}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{geom\_histogram}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x=}\NormalTok{biasesNull, }\AttributeTok{y =}\NormalTok{ ..density..), }\AttributeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\AttributeTok{bins=}\DecValTok{50}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme\_tufte}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{labs}\NormalTok{(}\AttributeTok{title=}\StringTok{"10k biases for identical means and sd =.05"}\NormalTok{)}\SpecialCharTok{+} \FunctionTok{xlab}\NormalTok{(}\StringTok{"bias"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-4-1} \end{center}
\normalsize

Now, let's simulate a situation where the means are identical but the
standard deviation is much smaller, .001.

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{124}\NormalTok{)}
\NormalTok{t1v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}
\NormalTok{t2v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}
\NormalTok{t3v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}
\NormalTok{t4v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-6-1} \end{center}

\noindent When you look at the datapoints, do you have the impression of
a strong bias here? We wouldn't think so. But now let's run the
calculations from (\textbf{Caliskan2017semanticsBiases?}).

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{factorV }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1v),}\FunctionTok{s}\NormalTok{(t2v),}\FunctionTok{s}\NormalTok{(t3v),}\FunctionTok{s}\NormalTok{(t4v)))}
\NormalTok{numeratorV }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1v),}\FunctionTok{s}\NormalTok{(t2v)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3v),}\FunctionTok{s}\NormalTok{(t4v))}
\FunctionTok{print}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{factor =}\NormalTok{ factorV,}\AttributeTok{numerator =}\NormalTok{ numeratorV, }\AttributeTok{bias =}\NormalTok{ numeratorV }\SpecialCharTok{/}\NormalTok{ factorV))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $factor
## [1] 0.0006402666
## 
## $numerator
## [1] -0.001237637
## 
## $bias
## [1] -1.933003
\end{verbatim}

\normalsize

While the numerator and the factors changed a lot, the bias actually
increased. One reason bias increases is that once the standard deviation
goes down, so does the factor used in the calculation of bias.

Again, to see whether this metric would provide us with meaningful
information, let's simulate 10000 drawings.

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{124}\NormalTok{)}

\NormalTok{biasesLowVariance }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\DecValTok{10000}\NormalTok{)\{}
\NormalTok{t1v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}
\NormalTok{t2v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}
\NormalTok{t3v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}
\NormalTok{t4v }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.001}\NormalTok{))}

\NormalTok{factorV }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1v),}\FunctionTok{s}\NormalTok{(t2v),}\FunctionTok{s}\NormalTok{(t3v),}\FunctionTok{s}\NormalTok{(t4v)))}

\NormalTok{numeratorV }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1v),}\FunctionTok{s}\NormalTok{(t2v)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3v),}\FunctionTok{s}\NormalTok{(t4v))}

\NormalTok{biasesLowVariance[i] }\OtherTok{\textless{}{-}}\NormalTok{ numeratorV }\SpecialCharTok{/}\NormalTok{ factorV}
\NormalTok{\}}
\FunctionTok{ggplot}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{geom\_histogram}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x=}\NormalTok{biasesLowVariance, }\AttributeTok{y =}\NormalTok{ ..density..), }\AttributeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\AttributeTok{bins=}\DecValTok{50}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme\_tufte}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{labs}\NormalTok{(}\AttributeTok{title=}\StringTok{"10k biases for identical means and sd =.001"}\NormalTok{)}\SpecialCharTok{+} \FunctionTok{xlab}\NormalTok{(}\StringTok{"bias"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-8-1} \end{center}
\normalsize

Again, not so informative. Now, let's try sampling from distributions
where there in fact is a difference in means. Terms from \(X\) are on
average .1 similar to \(A\) (and still \(0\) to \(B\)), while terms from
\(Y\) are \(.1\) similar to \(B\) (and 0 to \(A\)). The standard
deviation is 0.05 in all the cases. There is a clear difference between
\(X\) and \(Y\) and quick visual inspection should tell us so.

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{766}\NormalTok{)}
\NormalTok{t1d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t2d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t3d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t4d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\normalsize

\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-10-1} \end{center}
\normalsize

\noindent  Is this clear difference mirrored in the calculations?

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{factorD2 }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d2),}\FunctionTok{s}\NormalTok{(t2d2),}\FunctionTok{s}\NormalTok{(t3d2),}\FunctionTok{s}\NormalTok{(t4d2)))}
\NormalTok{numeratorD2 }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d2),}\FunctionTok{s}\NormalTok{(t2d2)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3d2),}\FunctionTok{s}\NormalTok{(t4d2))}
\NormalTok{biasD2 }\OtherTok{\textless{}{-}}\NormalTok{ numeratorD2 }\SpecialCharTok{/}\NormalTok{ factorD2}
\NormalTok{biasD2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.490014
\end{verbatim}

\normalsize

\noindent The absolute value of the effect size is smaller than in the
null case with the same standard deviation. Let's simulate 10000
drawings:

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{biasesD2 }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\DecValTok{10000}\NormalTok{)\{}
\NormalTok{t1d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t2d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t3d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}
\NormalTok{t4d2 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.05}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.05}\NormalTok{))}

\NormalTok{factorD2 }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d2),}\FunctionTok{s}\NormalTok{(t2d2),}\FunctionTok{s}\NormalTok{(t3d2),}\FunctionTok{s}\NormalTok{(t4d2)))}
\NormalTok{numeratorD2 }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d2),}\FunctionTok{s}\NormalTok{(t2d2)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3d2),}\FunctionTok{s}\NormalTok{(t4d2))}
\NormalTok{biasesD2[i] }\OtherTok{\textless{}{-}}\NormalTok{ numeratorD2}\SpecialCharTok{/}\NormalTok{factorD2}
\NormalTok{\}}

\FunctionTok{ggplot}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{geom\_histogram}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x=}\NormalTok{biasesD2, }\AttributeTok{y =}\NormalTok{ ..density..), }\AttributeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\AttributeTok{bins=}\DecValTok{50}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme\_tufte}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{labs}\NormalTok{(}\AttributeTok{title=}\StringTok{"10k biases for different means and sd =.05"}\NormalTok{)}\SpecialCharTok{+} \FunctionTok{xlab}\NormalTok{(}\StringTok{"bias"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-12-1} \end{center}
\normalsize

\noindent Now suppose we keep the means the same but increase the
standard deviation to .1.

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{766}\NormalTok{)}
\NormalTok{t1d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t2d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t3d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t4d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\normalsize

\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-14-1} \end{center}
\normalsize

\noindent  Is this clear difference mirrored in the calculations?

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{factorD1 }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d1),}\FunctionTok{s}\NormalTok{(t2d1),}\FunctionTok{s}\NormalTok{(t3d1),}\FunctionTok{s}\NormalTok{(t4d1)))}
\NormalTok{numeratorD1 }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d1),}\FunctionTok{s}\NormalTok{(t2d1)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3d1),}\FunctionTok{s}\NormalTok{(t4d1))}
\NormalTok{biasD1 }\OtherTok{\textless{}{-}}\NormalTok{ numeratorD1 }\SpecialCharTok{/}\NormalTok{ factorD1}
\NormalTok{biasD1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.223023
\end{verbatim}

\normalsize

\noindent The absolute value of the effect size is smaller than in the
previous case!

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{biasesD1 }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{(}\DecValTok{10000}\NormalTok{)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\DecValTok{10000}\NormalTok{)\{}
\NormalTok{t1d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t2d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t3d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}
\NormalTok{t4d1 }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{A  =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{), }\AttributeTok{B =} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{5}\NormalTok{,.}\DecValTok{1}\NormalTok{,}\FloatTok{0.1}\NormalTok{))}

\NormalTok{factorD1 }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d1),}\FunctionTok{s}\NormalTok{(t2d1),}\FunctionTok{s}\NormalTok{(t3d1),}\FunctionTok{s}\NormalTok{(t4d1)))}
\NormalTok{numeratorD1 }\OtherTok{\textless{}{-}}  \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t1d1),}\FunctionTok{s}\NormalTok{(t2d1)) }\SpecialCharTok{{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{s}\NormalTok{(t3d1),}\FunctionTok{s}\NormalTok{(t4d1))}
\NormalTok{biasesD1[i] }\OtherTok{\textless{}{-}}\NormalTok{ numeratorD1}\SpecialCharTok{/}\NormalTok{factorD1}
\NormalTok{\}}

\FunctionTok{ggplot}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{geom\_histogram}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x=}\NormalTok{biasesD1, }\AttributeTok{y =}\NormalTok{ ..density..), }\AttributeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\AttributeTok{bins=}\DecValTok{50}\NormalTok{)}\SpecialCharTok{+}
  \FunctionTok{theme\_tufte}\NormalTok{()}\SpecialCharTok{+}\FunctionTok{labs}\NormalTok{(}\AttributeTok{title=}\StringTok{"10k biases for different means and sd =.001"}\NormalTok{)}\SpecialCharTok{+} \FunctionTok{xlab}\NormalTok{(}\StringTok{"bias"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/unnamed-chunk-16-1} \end{center}
\normalsize

This is is a bit better, but still quite some uncertainty is involved,
far from what systematically low mean-based p-values reported in the
papers might suggest. Of course, this is a bit of a caricature, as our
word lists were short (four protected words and 10 attributes). But the
word lists used in the actually papers are not much longer. In such a
set-up the key observations are as follows:

\begin{itemize}
\item
  Seemingly high bias measures might arise even if the underlying
  processes actually have the same mean.
\item
  Even if the mean remains the same, non-negligible changes can result
  from a shift in the standard deviation of the original process, and
  the change might go in the opposite direction than a visualisation of
  datapoints might suggest, because with the decrease of standard
  deviation, the factor decreases and the bias increases.
\item
  Even if the underlying means are the same, but the variation is
  different, the bias metric in the long run could tend toward a
  different value:
\item
  The lack of control group in the paper and our analysis indicates that
  without neutral baseline it is difficult to interpret the
  effectiveness of the metric.
\end{itemize}

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{mean}\NormalTok{(biasesD1); }\FunctionTok{mean}\NormalTok{(biasesD2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.563319
\end{verbatim}

\begin{verbatim}
## [1] 1.690664
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{median}\NormalTok{(biasesD1); }\FunctionTok{median}\NormalTok{(biasesD2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.638136
\end{verbatim}

\begin{verbatim}
## [1] 1.707511
\end{verbatim}

\normalsize

and so the point estimations of bias are sensitive to factors other than
the underlying process means.

\begin{itemize}
\item
  Even if there is a difference in means, the bias metric can be lower,
  and the uncertainty about it needs to be gauged.
\item
  The uncertainty resulting from including the raw datapoint variance
  into considerations is more extensive than the one suggested by the
  low p-values obtained from taking means as datapoints.
\end{itemize}

\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{quantile}\NormalTok{(biasesD2, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.275}\NormalTok{,}\FloatTok{0.975}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    27.5%    97.5% 
## 1.539149 2.166488
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{quantile}\NormalTok{(biasesD1, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.275}\NormalTok{,}\FloatTok{0.975}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    27.5%    97.5% 
## 1.300648 2.356815
\end{verbatim}

\normalsize

\hypertarget{bayesian-estimation}{%
\section{Bayesian estimation}\label{bayesian-estimation}}

One thing to notice is that in Manzini, Lim, Tsvetkov, \& Black (2019)
words from all three subclasses (in the case of religion by a subclass
we mean Christianity, Judaism, and Islam) were compared against all of
the stereotypes, which means that there was no distinction between cases
in which the stereotype is associated with a given sub-class, as opposed
to the situation in which it is associated with another one. Not all of
the stereotypical words have to be considered as harmful for all of the
sub-groups. One should investigate the sub-groups separately as some of
them may have stronger harmful associations than others.

We decided to add control groups in the form of two classes --- neutral
words and human-related words. Without a proper control group it is
quite hard to compare the resulting cosine distances and decide on their
significance in bias detection. We prepared approximately 230 neutral
words to double-check the prima-facie neutral hypothesis that their
cosine similarity to the protected words will oscillate around 0 (that
is, the distances will be around 1). This provides us with a more
reliable point of reference. Moreover, we added human attributes that
are associated with people in general to investigate whether the smaller
cosine distance between protected words and stereotypes can result
simply from the fact that the stereotype predicates are associated with
humans.

One may consider improving the bias detection method presented in
Manzini, Lim, Tsvetkov, \& Black (2019). As the chosen protected words
and attributes do not represent the whole biased population but only a
sample, it is justified to introduce the uncertainty measure to the
method. We decided to use Bayesian estimation to make the cosine
distance method more methodologically correct.

Our prior distributions for the Bayesian model reflect our first stab at
what we know or expect before we see the data. All our distributions are
fairly weak. The distribution for mean distance (cosine distance) is set
to fit (almost completely) between 0 and 2 with the mean equal to 1, in
line with the idea that most of the cosine distances may be considered
neutral and therefore their similarity will be around 0. That is, we
take it to be DISTRIBUTION. The prior distribution for standard
deviation is half-cauchy between 0 and 1, but unlikely to be as large as
1. Finally the prior distributions for the coefficients have most of
their mass between -1 and 1, with mean around 0. As we want to put
ourselves in skeptical shoes looking at Manzini, Lim, Tsvetkov, \& Black
(2019) and asking whether the data would indeed convince us of the
existence of bias, we decided to use slightly skeptical (mean 0), but
fairly weak (sd = ) as our regularizing priors. The model parameters
will have a posterior distribution obtained using either Hamiltionian
Monte Carlo methods (STAN) available through the rethinking package.

\begin{center}
\begin{figure}[!htb]\centering
   \begin{minipage}{0.55\textwidth}
  \includegraphics[width=7cm]{../images/religionCoeffs.jpeg}
   \end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/religionGoogleCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/genderCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/genderGoogleCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/raceCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/raceGoogleCoeffs.jpeg}
   \end{minipage}
\end{figure}

\end{center}

Let's analyse dataset-level coefficients together with their highest
posterior density intervals, for the three topic groups. We compare the
results for two different word emebeddings corpuses - RedditL2 and
GoogleNews. For Google embeddings the HPDIs for all class coefficients
(associated, different, human, and none) include zero. This can lead one
to a conclusion that the impact for associated, different, human and
neutral attribute is, when averaged, quite similar. This indicates how
including the uncertainty may change the use and interpretation of
Manzini, Lim, Tsvetkov, \& Black (2019) MAC metric. It seems that if one
focuses only on differences between means of means, this is too
simplistic.

In the case of Reddit word embeddings the situation is similar although
HPDI is below 0 for Gender class when looking at \texttt{associated} and
\texttt{different} mean coefficient. This can suggest that there is
indeed slightly stronger impact of these attributes. One should also
notice how in general associated and different coefficients have quite
similar HPDI range, the highest observed absolute difference is equal to
only 0.1. This suggests that again the impact of associated attributes
and different ones is not clear at first sight. Finally, let's compare
the HPDIs for Reddit and Reddit debiased datasets. For Religion and Race
dataset there is a minor shift (in absolute values the highest change is
equal to approximately 0.1) of the mean coefficients towards zero.
However for Gender dataset there is no significant change. This is of
course a general look at the group coefficients. Let's now analyze the
individual words.

Looking at Reddit individual words one can observe that for Religion the
cosine distance results for the associated attributes are for most of
the words close to human attributes as well. This can suggest that in
some cases words concerning humans can have higher similarity with some
protected words independently of whether they are stereotypical or
neutral-human words. One should also pay attention to the fact that for
all of the associated and different attributes in Religion, the
uncertainty intervals overlap at some point. What is even more
surprising is that for protected words, such as ``torah'' the associated
attribute has the the cosine distance slightly over 1, which means no
positive similarity! If the protected words that we chose do not have
high similarity with harmful associated attributes, then one should
consider at least three scenarios. The first one is that the choice of
the protected words and attributes may be corrupted. The second one is
that the metric is unable to catch the hidden bias properly. The third
one is that there is actually no bias between the words. Regardless of
which scenario one considers, it is essential to take a look at the
individual values before averaging them or aggregating in other ways. It
seems that using Bayesian method can enhance the process of verifying
the hypotheses concerning the choice of protected words and attributes.

Surprisingly in Gender one can observe high cosine similarity values
between some female stereotypical professions and male protected words.
If a word stereotypically associated with females has low cosine
distance to male protected words, then one should investigate the issue
further. The reasons for this unexpected cosine distance may lie in the
frequency of appearance of the protected words in the raw data. Some of
the groups (like Muslim people or females) may have less representation
in the data that is taken as an input for word embeddings. Therefore, if
we assume that the MAC detects actually co-occurrence only, it makes
sense that they can have high similarity with associated attributes and
lower similarity with different ones. At the same time male protected
words and other religions can have high similarity with different
attributes because they have greater representation in the dataset in
general and occur close to much more concepts. Cosine distance seems to
capture the information on the co-occurrence of words and not on the
semantic similarity strictly speaking.

In Google, one may notice interesting phenomena as well. Although the
GoogleNews dataset is larger, trained on different data sources and with
the use of more dimensions, some of the results are similar to the ones
obtained for the Reddit corpus. For Religion almost all of the values
for associated class intersect with different class as well. In the case
of Race it is 100\% of the available words. This indicates again that it
is not clear how the metric should be used. Quite a different situation
takes place for Gender, where the similarity between associated and
different is present mostly for the male protected words. This means
that male words have high similarity with both male stereotypical
professions and female ones. However, for females, the similarity is
high mostly only for the female stereotypical professions. This
observation could not be made when using MAC metric only, as it requires
investigating the individual words, and providing uncertainty.

\hypertarget{effects-of-debiasing}{%
\section{Effects of debiasing}\label{effects-of-debiasing}}

One may verify if the dataset-level coefficients changed after the
debiasing method used in Manzini, Lim, Tsvetkov, \& Black (2019). In
some cases there is a minor change to the right, the mean value is
closer to 0. The \texttt{associated} and \texttt{differen} attributes
are now more overlapping than before. However, the uncertainty intervals
are still high and therefore the significance of the change cannot be
considered as enormous. The surprising obervation is that in the case of
\texttt{Gender} and \texttt{Race} groups, the neutral words shifted to
the left, further from the neutral position.

\begin{center}
\begin{figure}[!htb]\centering
   \begin{minipage}{0.55\textwidth}
  \includegraphics[width=7cm]{../images/religionCoeffs.jpeg}
   \end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/debiasedReligionRedditCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/genderCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/debiasedGenderRedditCoeffs.jpeg}
   \end{minipage}

  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7cm]{../images/raceCoeffs.jpeg}
\end{minipage}
   \begin {minipage}{0.43\textwidth}
    \includegraphics[width=7cm]{../images/debiasedRaceRedditCoeffs.jpeg}
   \end{minipage}
\end{figure}

\end{center}

When analyzing the Reddit debiased results one should investigate the
change of the individual cosine distance values as well. It seems that
not all of the protected words are treated equally when performing
debiasing. In religion, the values for associated class moved towards 0
for most of the words except for the ones for Islam, where still the
associated class has quite lower cosine distance than the different one.
Similar situation takes place in Gender, where female words have still
much lower cosine distance values for associated class than for the
different one. In the case of male protected words it is mostly almost
the same interval for associated and different. As the Gender data is
quite specific as the attributes are not harmful adjectives but (in an
ideal world) neutral professions, the aim (if we follow MAC assumptions)
should actually be to make the cosine distance same for both female and
male protected words. However, as we pointed out the situation after
debiasing can sometimes be better only for one protected group, which is
not the desired outcome. In the case of Religion, even after debiasing,
all of the protected words have intersections between associated and
different class and similarity greater than 0. As all of the attributes
in religion data are negative and harmful stereotypes, one should not
aim at making the distance between protected words and associated, and
different class the same but rather to move it towards 1.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\noindent \huge  \textbf{Appendix} \normalsize

\noindent \Large \textbf{Word lists, including human and neutral predicates}
\normalsize

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\vspace{-3mm}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Manzini2019blackToCriminal}{}%
Manzini, T., Lim, Y. C., Tsvetkov, Y., \& Black, A. W. (2019).
\emph{Black is to criminal as caucasian is to police: Detecting and
removing multiclass bias in word embeddings}. Retrieved from
\url{http://arxiv.org/abs/1904.04047}

\leavevmode\hypertarget{ref-Venkat2018Curse}{}%
Venkat, N. (2018). \emph{The curse of dimensionality: Inside out}.
\url{https://doi.org/10.13140/RG.2.2.29631.36006}

\end{CSLReferences}

\end{document}