### 2021-07-27 presentationESSLLI21/presentationESSLLI.Rmd
added presentation template

---
title: "\\large A bayesian method of cosine-based word2vec bias estimation"
author: "Alicja Dobrzeniecka \\& Rafal  Urbaniak \\footnotesize \\newline (LoPSE research group, University of  Gdansk)" 
date:  "ExpSem2021, ESSLLI"
fontsize: 10pt
output:  
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
classoption: x11names, dvipsnames, bibspacing,natbib
urlcolor: blue
bibliography: [../references/bibliographyRafal.bib]
csl: [../references/apa-6th-edition.csl] 
link-citations: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(ggExtra)
library(ggthemes)
library(latex2exp)
th <- theme_tufte(base_size = 20)
options(kableExtra.latex.load_packages = FALSE)
```

\frametitle{The gist of the project}

\pause

### Classical probabilism in epistemology (and in applications)

Represent uncertainty by a single probabilistic measure (point estimates).

\vspace{-3mm}

\begin{align*}
\pr(\mathsf{recovery}\vert \mathsf{treatment}) & = .6\\
\end{align*}

\vspace{-6mm}

\pause

### Imprecise probabilism

Evidence (e.g. expert evidence) sometimes fails to warrant such precision.

\vspace{-3mm}

\begin{align*}
\pr(\mathsf{recovery}\vert \mathsf{treatment}) & \in [.4,.6]\\
\end{align*}

\vspace{-6mm}

\pause

### Epistemological difficulties of imprecise probabilism

- hard to model the weight of and combine evidence.

- inertia of agnosticism.

- growth of uncertainty  in light of new evidence.

## The gist of the project

### My novel contribution

- Meet the desiderata motivating imprecision.

- Avoid these difficulties.

- Do so by using higher-order probabilities.

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%"}
p <- ggplot(data = data.frame(probability = seq(0,1, by = .25)),mapping = aes(x = probability))+th +ylab("density")+scale_x_continuous(breaks = seq(0,1, by = .1))

agentPrior <- function(p) dbeta(p,28,28)

priorPlot <- p + stat_function(fun=agentPrior)+ggtitle("Uncertainty about recovery")+xlab("probability of recovery if treated")

priorPlot
```

## Motivations for imprecision  \large (straighforward considerations)

\pause

### Radical lack of information

$\pr(\mathsf{It \,\, will \,\,  snow\,\,  in\,\,  Boston\,\,  on \,\, New\,\,  Year's\,\,  Eve\,\,  2026}) =$ ?

\pause

<!-- ### Mere bounds in reasoning -->

<!-- - Say $\pr(A) = .5$, $\pr(B) = .3$, but you don't know if they're independent. -->

<!-- \vspace{-2mm} -->

<!-- - We can only calculate bounds for $\pr(A \et B)$, $\pr(A \vee B)$, etc. -->

<!-- \pause  -->

### Weight of evidence

- You've seen 4/10 patients recover. $\pr(\mathsf{recovery})=.4$?

\pause
\vspace{-2mm}

- You've seen 40/100 patients recover. $\pr(\mathsf{recovery})=.4$?

\footnotesize (You could try to use classical statistical tools, such as confidence intervals,\linebreak  but they have their own problems.)

## Motivations for imprecision \large (uniformity and agnosticism)

### Agnosticism preservation requirement

 Agnosticism about uniquely connected variables should be the same.

\vspace{-5mm}

 \[X \mbox{ and } Y=X+2 \mbox{ or } Z = X^2 \,\,\, \mbox{ for $10>X>0$.}\]

\pause

### The trouble

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%"}
p <- ggplot(data = data.frame(value = 0:10), 
            mapping = aes(x = value))

unifX <-  function(value) punif(value,min = 0, max = 10)

p +   stat_function(fun=unifX)+
  ylab("density")+theme_tufte(base_size = 25)+labs(title = "Uniform cumulative probability for X in (0,10)")+xlab("X")+ylab("Cumulative probability")+scale_x_continuous(breaks = 0:10)+
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue")+
  geom_segment(aes(x = 0, y = .5, xend = 5, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue")

```

\vspace{-9mm}

\begin{align*}\pr(X<5) = .5 = \pr(X^2 < 25)
\end{align*}

## Motivations for imprecision \large (uniformity and agnosticism)

### The trouble

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", message = FALSE, warning = FALSE}
p <- ggplot(data = data.frame(value = 0:10), 
            mapping = aes(x = value))

unifY <-  function(value) punif(sqrt(value),min = 0, max = 10)

p +   stat_function(fun=unifY)+
  ylab("cumulative probability")+theme_tufte(base_size = 25)+
  labs(title = TeX("Resulting cumulative probability for $X^2$"))+scale_x_continuous(breaks = 0:100)+xlab(TeX("X^2"))+xlim(c(0,100))+
  geom_segment(aes(x = 25, y = 0, xend = 25, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue")+
  geom_segment(aes(x = 0, y = .5, xend = 25, yend = .5), lty = 2, alpha = .2, size = .4, color = "skyblue")

```

## Motivations for imprecision \large (expert evidence)

### Weather forecasters

\vspace{-1mm}

\begin{tabular}{ll}
Al: & $\pr(\mathsf{It \,\,  will\,\,  rain\,\,  tomorrow})=.6$\\
Bert: & $\pr(\mathsf{It \,\,  will\,\,  rain\,\,  tomorrow})=.8$
\end{tabular}

\pause

### Linear pooling

\vspace{-4mm}

\begin{align*}
\pr(\mathsf{It \,\,  will\,\,  rain\,\,  tomorrow}) & = 
\alpha \times .6 + \beta \times .8 \,\,\,\,\,  (where \,\, \alpha + \beta = 1.)
\end{align*}

\pause

### Mathematical impossibilities

- Experts's agreement on independence is not preserved by linear pooling. [@List2011Group]

- You can't at the same time hold the following: [@Gallow2018No]

\vspace{-7mm}

\begin{align}
\pr(A=B) & <1\\
\pr(r\vert A=a) & = a \\
\pr(r\vert B=b) & = b \\
\forall a,b \, \pr(r \vert A =a, B = b) & =  \alpha a + \beta b
\end{align}

## Imprecise probabilism and the motivations

### The main claim

\justify Represent uncertainty by  a set of probability measures.

\vspace{1mm}
\scriptsize

[@Levi1974;@Grdenfors1982;@Jeffrey1983-JEFBWA-2;@walley1991statistical;@Kaplan1996;@Joyce2005;@Grove2009;@elkin2017imprecise]

\pause

### Snowing in Boston?

\vspace{-1mm}

Say, (.5,1).

\vspace{1mm}
\scriptsize

[@Mahtani2019Imprecise]

\pause

### Weight of evidence?

\vspace{-1mm}

The more evidence you have, the more narrow the interval.

\vspace{1mm}
\scriptsize

[@Kaplan1996;@Joyce2005;@Sturgeon2008-STURAT]

\pause

<!-- ## Imprecise probabilism and the motivations -->

### Agnosticisim?

\vspace{-1mm}

[0,1] interval instead of a single uniform distribution.

\pause

### Multiple sources?

\vspace{-1mm}

Just put all those probabilities in one set.

## Epistemological challenges

\pause

### Radical uncertainty and weight

Say one experiment observed 40/100  and  another 8/15 recoveries.

Simply putting the frequencies in a set won't do justice to the evidence.

\scriptsize 
\vspace{1mm}

[@Bradley2019Imprecise]

\pause

### Dilation

  \justify Whenever your roomate goes to the track he flips a fair coin. If it comes up heads he bets on  Speedy. Otherwise he places no bets. He comes home from the track grinning.  You're now more uncertain about  $\pr(\mathsf{heads})$ than before.

\scriptsize

\vspace{1mm}
 [@walley1991statistical;@seidenfeld1993;@White2009-WHIESA;@Topey2012;@pedersen2014demystifying;@Hart2015;@Bradley2016]

## Epistemological challenges

### Belief inertia

Say you're fully agnostic about $H$: $P(H) = [0,1]$.

\begin{center}
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", warning=FALSE, message = FALSE, dpi = 800}
p <- ggplot(data = data.frame(value = 1:7), 
            mapping = aes(x = value))+ theme_tufte(base_size = 24)+
  theme(axis.line.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank())+
  ylab("probability")+xlim(c(1.8,6.2))

p + geom_segment(aes(x = 2, y = 0, xend = 2, yend = 1), 
                 lineend = "square")+
   geom_segment(aes(x = 6, y = 0, xend = 6, yend = 1), 
                 lineend = "square")+
   annotate("text", x = 1.8, y = 0, label = TeX("$p_F$"), size = 7 )+
  annotate("text", x = 1.8, y = 1, label = TeX("$p_T$"), size = 7)+
  annotate("text", x = 1.8, y = .4, label = TeX("$p_{1}$"), size = 7)+
  annotate("text", x = 1.8, y = .6, label = TeX("$p_{2}$"), size = 7)+
  geom_segment(aes(x = 1.95, y = 0, xend = 2.05, yend = 0), 
             lineend = "square")+
  geom_segment(aes(x = 1.95, y = 0.4, xend = 2.05, yend = 0.4), 
               lineend = "square")+
  geom_segment(aes(x = 1.95, y = 0.6, xend = 2.05, yend = 0.6), 
               lineend = "square")+
  geom_segment(aes(x = 1.95, y = 1, xend = 2.05, yend = 1), 
               lineend = "square") + #now the second one
  annotate("text", x = 6.2, y = 0, label = TeX("$p_F'$"), size = 7)+
  annotate("text", x = 6.2, y = 1, label = TeX("$p_T'$"), size = 7)+
  annotate("text", x = 6.2, y = .2, label = TeX("$p_{1}'$"), size = 7)+
  annotate("text", x = 6.2, y = .5, label = TeX("$p_{2}'$"), size = 7)+
  geom_segment(aes(x = 5.95, y = 0, xend = 6.05, yend = 0), 
               lineend = "square")+
  geom_segment(aes(x = 5.95, y = 0.2, xend = 6.05, yend = 0.2), 
               lineend = "square")+
  geom_segment(aes(x = 5.95, y = 0.5, xend = 6.05, yend = 0.5), 
               lineend = "square")+
  geom_segment(aes(x = 5.95, y = 1, xend = 6.05, yend = 1), 
               lineend = "square")+ 
                 geom_segment(aes(x = 2.1, xend = 5.9, y = 0,
                yend = 0), arrow = arrow(length = unit(0.08, "inches")),
                color = "red", size = 0.6)+ 
  geom_segment(aes(x = 2.1, xend = 5.9, y = 1,
                   yend = 1), arrow = arrow(length = unit(0.08, "inches")),
               color = "red", size = 0.6)+ 
  geom_segment(aes(x = 2.1, xend = 5.9, y = 0.4,
                   yend = .2), arrow = arrow(length = unit(0.08, "inches")),
               color = "red", size = 0.6)+ 
  geom_segment(aes(x = 2.1, xend = 5.9, y = 0.6,
                   yend = .5), arrow = arrow(length = unit(0.08, "inches")),
               color = "red", size = 0.6)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.1, yend = 0.06), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.2, yend = 0.1), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.3, yend = 0.14), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.45, yend = 0.28), arrow = arrow(length = unit(0.06, "inches")),
                                                            color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.5, yend = 0.36), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.55, yend = 0.45), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.65, yend = 0.55), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.7, yend = 0.65), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.75, yend = 0.72), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.79, yend = 0.8), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.84, yend = 0.86), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)+ 
  geom_segment(aes(x = 2.2, xend = 5.8, y = 0.9, yend = 0.93), arrow = arrow(length = unit(0.06, "inches")),
               color = "skyblue", size = 0.3)
```
\end{center}

<!-- \begin{center} -->
<!-- \includegraphics[width = 6cm]{inertia.jpg} -->
<!-- \end{center} -->

Then $\pr(H \vert E) =[0,1]$, no matter what the evidence says.

\vspace{1mm}
\scriptsize

[@Joyce2010,@Dodd2012;@Rinard2013;@Bradley2014;@Vallinder2017;@Moss2020]

## Epistemological challenges

### Mystery coins

- A fair draw of one of two biased coins, with biases $.4, and .6$.

\pause

\vspace{-2mm}

- Imprecise reaction: $\{.4, .6\}$.

\pause

\vspace{-2mm}

- Precise probabilism: expected bias of $.5$.

\pause

### Question

As far as accuracy is concerned, why prefer the imprecise way?

Say the actual bias is .4. A first stab:

\vspace{-3mm}

\begin{align*}
\mathsf{inaccuracy(imprecise)} & = \frac{0 + .2}{2} = .1\\
\mathsf{inaccuracy(precise)} & =  .1\\
\end{align*}

\vspace{-5mm}
\scriptsize

[@Seidenfeld2012forecasting; @MayoWilson2016; @schoenfield2017AccuracyRationalityImprecise]

## Higher-order approach

### The mystery coin with  $(.2,.8)$ and 1 head in five tosses

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%"}
theta <- c(.4, .6) #possible parameter values
priorTheta <- c(.2,.8) #priors assigned to them
data <- c(1,0,0,0,0) #1 heads with five tosses
z = sum(data) # number of 1's in Data
N = length(data) #number of tosses

# Compute the Bernoulli likelihood at each value of Theta:
pDataIfTheta <- dbinom(z,N,theta)
# Compute the probability of th evidence and 
# the posterior via Bayes' rule:
pData <- sum(pDataIfTheta * priorTheta)
pThetaIfData <- pDataIfTheta * priorTheta / pData

#put together
table <- data.frame(theta,  priorTheta,
                     pData,  pDataIfTheta, pThetaIfData)

#plot of the prior
prior <- ggplot(table, aes(x = theta, y = priorTheta))+  
  geom_bar(stat="identity", alpha = .8)+ 
  scale_x_continuous(breaks = theta)+th+ylim(c(0,1))+ggtitle("Prior")

#plot of the likelihood
likelihood <-  ggplot(table, aes(x = theta, y = pDataIfTheta))+
  geom_bar(stat="identity", alpha = .8)+ 
  scale_x_continuous(breaks = theta)+
  th+ylim(c(0,1))+ggtitle("Likelihood")

#plot of the posterior
posterior <- ggplot(table, aes(x = theta, y = pThetaIfData))+  #set up
  geom_bar(stat="identity", alpha = .8)+  #barplot layer
  scale_x_continuous(breaks = theta)+
  th+ylim(c(0,1))+ggtitle("Posterior")

#convert the table to graphics, transpose for readability
tableGG <- ggtexttable(t(round(table,3)))

#plot together
ggarrange(prior, likelihood, posterior, tableGG)
```

## Higher-order approach

### Two experts, uniform prior

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%"}
p <- ggplot(data = data.frame(probability = seq(0,1, by = .25)),mapping = aes(x = probability))+th +ylab("density")+scale_x_continuous(breaks = seq(0,1, by = .25))+ylim(c(0,10))

agentPrior <- function(p) dbeta(p,1,1) 
al <- function(p) dbeta(p,5,20) 
bert <- function(p) dbeta(p,8,12) 
posterior <- function(p) dbeta(p,19,62)

priorPlot <- p + stat_function(fun=agentPrior)+ggtitle("Agent\'s prior")
alPlot <- p + stat_function(fun=al)+ggtitle("If Al says .2")

bertPlot <- p + stat_function(fun=bert)+ggtitle("If Bert says .4")
posterior <- p + stat_function(fun=posterior)+ggtitle("Agent's posterior")

ggarrange(priorPlot, alPlot, bertPlot, posterior)
```

## The key conjecture

### With higher-order methods, it will be possible to:

- match the representation of uncertainty with the evidence,

- combine various sources of uncertainty without running into the known problems,

- use existing mathematics to make sense of accuracy in such contexts,

- capture reasoning with radical uncertainty without dilation and belief inertia.

## Tasks (discussion)

\footnotesize

- Relate this representation to weight of evidence.

\vspace{-2mm}

- Go beyond grid approximation and beta representation.

- Represent  agnosticism so that:

    - agnosticism  preservation holds, and

    - inertia fails.

\vspace{-2mm}

- Represent opinion aggregation more generally, compare to known limitations and features of already studied opinion pooling methods.

\vspace{-2mm}

- Investigate dilation-like phenomena in this context.

\vspace{-2mm}

- Study known distance functions between probabilistic measures as candidates for key elements of accuracy measures.

\vspace{-2mm}

- Explain away the apparent contrast between evidence-grounding and accuracy considerations [@easwaranAndFitelson2015].

\begin{center}
\Large Thank you for your attention!
\end{center}

## References
\tiny