### 2022-03-08 presentations/presentationBoston/presentationBoston.Rmd
presentation Boston added cosine etc.

---
title: '\Large Taking uncertainty seriously  \newline \normalsize A  Bayesian approach
  to  word embedding bias estimation '
author: Alicja Dobrzeniecka \& Rafal  Urbaniak \footnotesize \newline (LoPSE research
  group, University of  Gdansk)
date: "Boston, April Fools' Day"
output:
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
  slidy_presentation: default
fontsize: 10pt
classoption: x11names, dvipsnames, bibspacing,natbib, table
urlcolor: blue
bibliography: ../../references/cosineReferences1.bib
csl: ../../references/apa-6th-edition.csl
link-citations: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(ggExtra)
library(ggthemes)
library(ggforce)
library(latex2exp)
th <- theme_tufte(base_size = 20)
options(kableExtra.latex.load_packages = FALSE)
```

## Presentation plan

- Bias in word embeddings
- WEAT and MAC methods
- Methodological problems
- Limitations of pre-averaging in bias detection methods
- Accounting for uncertainty with Bayesian approach

## Word2vec

### Question

How to sensibly represent words with numbers?

\pause

### One-hot encoding

Well, you could use 30k binary vectors with a slot for each lexical unit\dots

\pause

\dots \dots but this would be inefficient and wouldn't capture any relations between words.

## Word2vec

\begin{center}
 \includegraphics[width = 9cm]{images/perceptron1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

\footnotesize

\begin{block}{Rosenblatt's perceptron}
\begin{itemize}
\item Inputs (pixel intensities) with weights
\item Nodes with activation levels from 0-1
\item (Perhaps) 0-1 output based on a threshold
\end{itemize}
\end{block}

## Word2vec

\begin{center}
 \includegraphics[width = 9cm]{images/perceptron1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

\footnotesize

\begin{block}{Learning}
\begin{itemize}
\item Start with random weights
\item Test on a case:
\begin{itemize}
\item If right, don't change weights.
\item If wrong, change weights a bit, with focus on the ones more responsible for the judgment:
\begin{align*}
w_j & \leftarrow w_j = \overbrace{\eta}^{\text{learning rate}}(\underbrace{t}_{\text{correct output}} - \overbrace{y}^{\text{actual output}})\underbrace{x_j}_{\text{actual input}}
\end{align*}
\end{itemize}
\end{itemize}

\end{block}

## Word2vec

\begin{columns}
\column{0.45\linewidth}

\begin{center}
 \includegraphics[height = 6cm, width = 5cm]{images/perceptron2.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

\column{0.5\linewidth}

\footnotesize

\begin{itemize}
\item Each hidden unit takes a weighted sum of 324 inputs and passes on its activation level as input to outer layer units.

\item Activation levels of outer layers are interpreted as network's levels of confidence in a classification problem.

\item Learning: back-propagation (gradient descent: approximate  the direction of steepest descent in the error surface w.r.t to weights, modify accordingly).
\end{itemize}

\end{columns}

## Word2vec

### Distributional semantics

\begin{itemize}
\item "You shall know a word by the company it keeps" (John Firth, 1957)

\item "the degree of semantic similarity between two linguistic expressions $A$ and $B$ is a function of the similarity of the linguistic contexts in which $A$ and $B$ can appear." (A. Lenci, 2008)

\end{itemize}

\pause

\begin{center}
 \includegraphics[height = 5cm, width = 5cm]{images/similarity1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

## Word2vec

### Google and Mikolov

\emph{Efficient Estimation of Word Representation in Vector Space}, 2013

Let's train a neural network and use vectors of weights!

\begin{center}
 \includegraphics[height = 5cm, width = 6cm]{images/word2vec1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

## Word2vec

### Google and Mikolov

\emph{Efficient Estimation of Word Representation in Vector Space}, 2013

Let's train a neural network and use vectors of weights!

\begin{center}
 \includegraphics[height = 3cm, width = 7cm]{images/word2vec2.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

## Word2vec

word | 1 | 2 | 3 | 4 | ... 
|---|---|---|---|---|---
woman | 0.456 | 0.267 | 0.675 | 0.131| ...
man | 0.451 | 0.897 | 0.472 | 0.088| ...

### Question

How is this supposed to capture semantic relations?

\pause

### General idea

Similarity in vector direction.

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine1,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Vectors}
\begin{align*}
a  & = [1,2]\\
b  &= [3,2]
\end{align*}

\end{block}
\pause

\begin{block}{Dot product}

\begin{align*}
a \cdot b & = a_1 b_1 + a_2 b_2\\
a \cdot a & = a_1^2 + a_2 ^ 2 \\
\lVert a\rVert & = \sqrt(a \cdot a)
\end{align*}

\end{block}

\end{columns}

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine2,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)
bma <- b - a

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)+
  geom_segment(aes(x = 0, y = 0, xend = bma[1], yend = bma[2]),
      arrow = arrow(length = unit(0.02, "npc")), color = "grey")+ 
  annotate(geom = "text", label  = "b - a = [2,.5]", x = 2.2, y = .9, size = 7)+
  theme_tufte(base_size = 24)

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Vectors}

\begin{align*}
a  & = [1,2]\\
b  &= [4,4]
\end{align*}

\end{block}

\begin{block}{Dot product}

\begin{align*}
a \cdot b & = a_1 b_1 + a_2 b_2\\
a \cdot a & = a_1^2 + a_2 ^ 2 \\
\lVert a\rVert & = \sqrt(a \cdot a)
\end{align*}

\end{block}

\begin{block}{Vector difference}

\begin{align*}
b - a & = [b_1- a_1, b_2 - a_2 ]
\end{align*}

\end{block}

\end{columns}

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine3,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)+
  geom_segment(aes(x = a[1], y = a[2], xend = b[1], yend = b[2]),
      arrow = arrow(length = unit(0.02, "npc")), color = "grey")+ 
  annotate(geom = "text", label  = "b - a = [2,.5] (shifted)", x = 2, y = 2.7, size = 7)+
  theme_tufte(base_size = 24)

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Vectors}

\begin{align*}
a  & = [1,2]\\
b  &= [4,4]
\end{align*}

\end{block}

\begin{block}{Dot product}

\begin{align*}
a \cdot b & = a_1 b_1 + a_2 b_2\\
a \cdot a & = a_1^2 + a_2 ^ 2 \\
\lVert a\rVert & = \sqrt(a \cdot a)
\end{align*}

\end{block}

\begin{block}{Vector difference}

\begin{align*}
b - a & = [b_1- a_1, b_2 - a_2 ]
\end{align*}

\end{block}

\end{columns}

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine4,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)
#atan2(a[1] , a[2])
#atan2(b[1] , b[2])

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)+
  geom_segment(aes(x = a[1], y = a[2], xend = b[1], yend = b[2]),
      arrow = arrow(length = unit(0.02, "npc")), color = "grey")+ 
  annotate(geom = "text", label  = "b - a = [2,.5] (shifted)", x = 2, y = 2.7, size = 7)+
  theme_tufte(base_size = 24)+
  annotate(geom = "text", label  = expression(theta), x = .6, y = .78, size = 7)+
  geom_arc(aes(x0 = 0, y0 = 0, r = 1.3, 
               start = .463, end = .876))

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Angle}

\begin{align*}
\lVert b - a \rVert^2 &  = \lVert  b \rVert^2 + \lVert a \rVert^2 - 
  2 \lVert b \rVert   \lvert  a \rVert  \cos \theta  \\
b \cdot a  & = \lVert b\rVert \lVert  a \rVert \cos \theta \\ 
\cos \theta & = \frac{b \cdot a}{\lVert b\rVert  \lVert  a \rVert }
\end{align*}

\end{block}

\pause

\begin{block}{Orthogonality}
\begin{align*}
\cos ( 90^{\circ}) & = 0 \\ 
\frac{b \cdot a}{\lVert  b\rVert \lVert a \rVert} & = 0\\ 
 b \cdot a & = 0  
\end{align*}
\end{block}

\end{columns}

## Cosine similarity \& distance

<!-- - Dot product for normalized vectors -->

<!-- often used in discussion of bias -->

\vspace{-4mm}

\begin{align} \tag{Sim}
\mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\lVert  A \rVert \,\lVert B \rVert}
\\
\tag{Distance}
\mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B)
\end{align}

-  Geometric interpretation: direction (not length)

-  $\mathsf{cosineDistance}\in (0, 2)$

- Naive interpretation: proximity corresponds to semantic similarity

<!-- - Note cosine distance fails to meet triangle inequality -->

## Cosine-based measures of bias

### The worry

Word embeddings can learn implicit harmful biases

\pause

### The basic intuition

Stereotypically connected words are cosine-close

## Cosine-based measures of bias

### A visual example

\footnotesize

- "feminine" occupations: "homemaker", "nurse", "receptionist", "librarian", etc.

- "masculine" occupations:  "maestro",  "captain", "architect", "boss", etc.

\normalsize

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%"}
genderProjections <-  read.csv("datasets/genderProjections.csv")[,-1]
colnames(genderProjections)[4] <- "Stereotype" 
levels(genderProjections$Stereotype) <- c("male occupation", "female occupation")
ggplot(genderProjections, aes(x=he,y=she, color = Stereotype, shape = Stereotype))+geom_point(size = 6, alpha = 0.75)+theme_tufte(base_size = 27)+labs(title ="GloVe  on Wikipedia 2014 and Gigaword 5th ed.")+scale_color_manual(values =c("orangered4","chartreuse4"))+ 
  theme(legend.position = c(0.8,.15)) +
    scale_shape_manual(values=c(15, 19)) 
```
\normalsize

<!-- ## Cosine-based measures of bias -->

<!-- ### Example: direct bias -->

<!-- - The gender bias of a word $w$ is its projection on the gender direction $\vec{w} \cdot (\overrightarrow{he} - \overrightarrow{she})$ -->

<!-- <!-- (the gender direction is the top principal compontent of ten gender pair difference vectors).  -->

<!-- -  Given the (ideally) gender neutral words $N$ and the gender direction $g$ the direct gender bias is: -->
<!-- <!-- ($c$ is a parameter determining how strict we want to be): -->

<!-- \vspace{-2mm} -->

<!-- \begin{align} -->
<!-- \mathsf{directBias_c(N,g)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},g)\vert^c}{\vert N \vert } -->
<!-- \end{align} -->

<!-- \footnotesize  -->

<!-- [@Bolukbasi2016Man] -->

## Cosine-based measures of bias

### Example: Word Embedding Association Test (WEAT)

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "55%"}
words <- c("man", "he", "boss", "architect","nurse","receptionist","woman","she")
x <- c(.1,.1,.5,.5,.7,.7,1.2,1.2)
y <- c(.4,.3,.3,.2,.5,.4,.4,.3)
weat <- data.frame(words,x,y)

ggplot(weat, aes(x=x,y=y, label = words))+geom_point()+geom_text(check_overlap = TRUE,
                                                                 nudge_y = 0.018, size = 7 )+
    geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.5, yend=.2), , linetype =2, arrow=arrow(), size=.05, color="blue") +
  geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.5, yend=.3), linetype =2, arrow=arrow(), size=.05, color="blue")+
  geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.7, yend=.4), linetype =2, arrow=arrow(), size=.05, color="orangered")+
  geom_segment(data=weat, mapping=aes(x=.1, y=.3, xend=.7, yend=.5), linetype =2, arrow=arrow(), size=.05, color="orangered")+
    annotate("text", x = c(.32,.32,.32,.32), y = c(.23,.29,.35,.39), 
             label = c(".7", ".6",".1", ".2") , color="black", 
             size=7,  fontface="bold")+
    annotate("text", x = c(.1,.5,.7,1.2), y = c(.55,.55,.55,.55), 
           label = c("X", "A","B", "Y") , color=c("black", "blue","orangered","black"), 
           size=10,  fontface="bold")+xlim(0,1.3)+
  theme_void(base_size = 25)
```
\normalsize

\pause

\footnotesize

- $s_1 = s(he,A,B) =  \frac{.6+.7}{2}  - \frac{.2+.1}{2} = .65-.15= .5$

- $s_2 = s(man,A,B) = .3$, \linebreak  $s_3 = s(woman,A,B) = -.6, s_4 = s(she, A, B) = -.3$

\vspace{-4mm}

\normalsize 
\begin{align*}
\mathsf{WEAT}(A,B) & = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93
\end{align*}

## Cosine-based measures of bias

### Example: Word Embedding Association Test (WEAT)

\begin{align*}
s(t,A,B) & = \frac{\sum_{a\in A}f(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}f(t,b)}{\vert B\vert}
\\
WEAT(A,B) & = \frac{
\mu\left(\{s(x,A,B)\}_{x\in X}\right) -\mu\left(\{s(y,A,B)\}_{y\in Y}\right) 
}{
\sigma\left(\{s(w,A,B)\}_{w\in X\cup Y}\right)
}
\end{align*}

-  $t$ is a term,  $A, B$ are sets of stereotype attribute words, $X$, $Y$ are protected group words

- For instance, $X$ might be a set of male names, $Y$ a set of female names, $A$ might contain stereotypically male-related career words, and $B$ stereotypically female-related family words

- $s$-values are used as datapoints in statistical significance tests

\footnotesize

\vspace{2mm}
[@Caliskan2017semanticsBiases] with extensions in [@Lauscher2019multidimensional] and applications in [@Garg2018years]

## Cosine-based measures of bias

### Our main target: Mean Average Cosine Similarity (MAC)

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "50%"}
words <- c("jew", "muslim", "christian", "terrorist","dirty","conservative","familiar","greedy","cheap")
x <- c(.1,.1,.1,.4,.4,.7,.7,1.1,1.1)
y <- c(.4,.3,.2,.1,.2,.3,.4,.1,.2)
mac <- data.frame(words,x)

ggplot(mac, aes(x=x,y=y, label = words))+geom_point()+geom_text(check_overlap = TRUE,
                                                                 nudge_y = 0.02, size = 9 )+
  geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.1), , linetype =2, arrow=arrow(), size=.03, color="blue")+geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.4, yend=.2), , linetype =4, arrow=arrow(), size=.0, color="blue") + geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.3), linetype =2, arrow=arrow(), size=.05, color="orangered")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=.7, yend=.4), linetype =2, arrow=arrow(), size=.05, color="orangered")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.1), linetype =3, arrow=arrow(), size=.05, color="chartreuse4")+ geom_segment(data=mac, mapping=aes(x=.1, y=.3, xend=1.1, yend=.2), linetype =2, arrow=arrow(), size=.05, color="chartreuse4")+ annotate("text", x = c(.1,.4,.7,1.2), y = c(.45,.45,.45,.45),  label = c("T", "A[1]","A[2]", "A[3]") ,parse = TRUE, color=c("black", "blue","orangered","chartreuse4"),            size=10,  fontface="bold")+xlim(0,1.3)+theme_void(base_size = 25)

```
\normalsize

\vspace{-4mm}

\footnotesize 
\begin{align*}
 s_1 & = s(muslim,A_1) = \frac{cos(muslim,dirty)+cos(muslim,terrorist)}{2}\\
s_2 & = s(muslim,A_2) = \frac{cos(muslim,familiar)+cos(muslim,conservative)}{2}\\
& \vdots \end{align*}

\vspace{-6mm}

\normalsize 
\begin{align*}
 MAC(T,A) & = \mathsf{mean}(\{s_i \vert i \in 1, \dots, k\})
\end{align*}

## Cosine-based measures of bias

### Our main target: Mean Average Cosine Similarity (MAC)

<!-- - more general, multi-class setting -->

\begin{align*}
S(t_i, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) \\
MAC(T,A) & = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j)
\end{align*}

- $T = \{t_1, \dots, t_k\}$ is a class of protected words

- each $A_j\in A$ is a set of attributes stereotypically associated with a protected word

<!-- - For each protected word $T$ and each attribute class, they first take the mean for this protected word and all attributes in a given attribute class, and then take the mean of thus obtained means for all the protected words.  -->

- The t-tests they employ are run on average cosines used to calculate MAC

\vspace{2mm}

\footnotesize

[@Manzini2019blackToCriminal]

## Cosine-based measures of bias

### Our main target: Mean Average Cosine Similarity (MAC)

\footnotesize 
```{r religionTableHeadEarly,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
religion <- read.csv("datasets/religionReddit.csv")[-1]
colnames(religion) <- c("protectedWord","wordToCompare","wordClass",
                        "cosineDistance","cosineSimilarity","connection")
religion$wordClass <- as.factor(religion$wordClass)
levels(religion$wordClass) <- c("christian","human","jewish","muslim","neutral")
religionSmall <- religion[religion$wordClass != "neutral" & religion$wordClass != "human" ,]
set.seed(127)
religionSmall <- religionSmall[sample(1:nrow(religionSmall),6),-c(3,6)]
rownames(religionSmall) <- c()
religionSmall  %>%  kable(format = "latex",booktabs=T,
                      linesep = "",  escape = FALSE, 
                      caption = "A few rows from the religion dataset") %>%
                      kable_styling(latex_options=c("scale_down"))
```
\normalsize

## Cosine-based measures of bias

### General challenges

- Gender-direction: insufficient indicator of
    bias  \footnotesize  [@Gonen2019lipstick]

\normalsize

- Use of analogies: unreliable \footnotesize  [@Nissim2020fair] \normalsize

- High sensitivity to irrelevant factors \footnotesize  [@zhang2020robustness] \normalsize

## Some methodological problems

###  Word list choice is unprincipled

- We run with it for comparison.

\pause

### No design considerations to sample size

- ETHAYARAJH IS YOUR criticizes WEAT uses Bernstein bounds and argues that we would need a bias specific dataset of size at least 11903 to claim that the system is biased (three times larger than WinoBias).

\pause

- We show progress can be made with more sensitive Bayesian methods.

### The form of the definition is suspicious

ETHAYARAJH UNDERSTANDING shows if there are two target words only WEAT is always maximal in one direction.

\pause

- We show the problem runs deeper and stems from pre-averaging, and we statistically gauge the uncertainty that arises from raw sample sizes.

## Some methodological problems

### No word class distinction and no control group

We make the subclasses clear, add human neutral predicates and neutral predicates for control. We used L2-Reddit corpus and GoogleNews (we present the results for Reddit for brevity).

\footnotesize 
```{r religionTableHeadLate,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
religion <- read.csv("datasets/religionReddit.csv")[-1]
colnames(religion) <- c("protectedWord","wordToCompare","wordClass",
                        "cosineDistance","cosineSimilarity","connection")
religion$wordClass <- as.factor(religion$wordClass)
levels(religion$wordClass) <- c("christian","human","jewish","muslim","neutral")
religionSmall <- religion[religion$wordClass != "neutral" & religion$wordClass != "human" ,]
religionControl <- religion[religion$wordClass == "neutral" | religion$wordClass == "human" ,]
set.seed(123154196)
religionDisplayA <- religionSmall[sample(1:nrow(religionSmall),4),]
religionDisplayB <- religionControl[sample(1:nrow(religionControl),4),]
religionDisplay <- rbind(religionDisplayA,religionDisplayB)
religionDisplay$cosineDistance <- round(religionDisplay$cosineDistance,3)
religionDisplay$cosineSimilarity <- round(religionDisplay$cosineSimilarity,3)
rownames(religionDisplay) <- c()
religionDisplay  %>%  kable(format = "latex",booktabs=T,
                      linesep = "",  escape = FALSE, 
                      caption = "Rows from extended religion dataset.") %>%
                      kable_styling(latex_options=c("scale_down"))
```
\normalsize

## Some methodological problems

### Outliers and surprisingly dissimilar words

We study those by  visualizations and uncertainty estimates.

## Some methodological problems

### Distances for  "muslim"

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%",message =FALSE, warning = FALSE}
library(tidyverse)
muslimWords <- c("imam","islam","mosque","muslim","quran")
muslim <- religion %>% filter(protectedWord %in% muslimWords)
muslimClass <- muslim %>% filter(protectedWord == "muslim")
neutralSample <- sample_n(filter(muslimClass,connection == "none"), 5)
humanSample <- sample_n(filter(muslimClass,connection == "human"), 5)
muslimVis <- muslimClass %>% filter(connection != "none" & connection !="human")
muslimVis <- rbind(muslimVis,neutralSample,humanSample)
#we plug in our visualisation script
source("../../functions/visualisationTools.R")
#two arguments: dataset and protected word
visualiseProtected(muslimVis,"muslim")+theme_void(base_size = 20)
```

## Some methodological problems

### Distances for  "priest"

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
library(tidyverse)
priestClass <- religion %>% filter(protectedWord == "priest")
neutralSample <- sample_n(filter(priestClass,connection == "none"), 5)
humanSample <- sample_n(filter(priestClass,connection == "human"), 5)
priestVis <- priestClass %>% filter(connection != "none" & connection !="human")
priestVis <- rbind(priestVis,neutralSample,humanSample)

#we plug in our visualisation script
source("../../functions/visualisationTools.R")
#two arguments: dataset and protected word
visualiseProtected(priestVis,"priest")+theme_void(base_size = 20)
```

## Some methodological problems

### No principled interpretation

Religion Debiasing  | MAC (distance)
------------- | -------------
Biased        | 0.859
Hard Debiased | 0.934
Soft Debiased ($\lambda$ = 0.2) | 0.894

- What values are sufficient for the presence of bias and what differences are sign of real improvement?

- Low $p$-values are not high effect indicators!

- We compare HPDIs.

<!--  What may attract attention is the fact that the value of cosine distance in "Biased" category is already quite high even before debiasing. High cosine distance indicates low cosine similarity between values. One could think that average cosine similarity equal to approximately 0.141 is not significant enough to consider it as bias. However the authors aim to mitigate "biases" in vectors with such great distance to make it even larger. Methodologically the question is, on what basis is this small similarity still considered as a proof of the presence of bias, and whether these small changes are meaningful. This is in general the problem of scale and the lack of universal intervals. In contrast, statistical intervals will help us decide whether a given cosine similarity is high enough to consider the words to be more similar than if we chose them at random. We will use highest posterior density intervals, in line with Bayesian methodology.  -->

## The problem with pre-averaging

### Key conceptual issues

- It throws away information about sample sizes
- It ignores variation in the raw data, which leads to false confidence

\pause

### Our simulations

Suppose all similarities for two classes are randomly drawn from the same distribution, $\mathsf{Normal}(\mu = 0, \sigma = .08)$, you still can get a really high WEAT!

## The problem with pre-averaging

### Simple case: two pws, four terms

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%"}
set.seed(123)
t1 <- data.frame(A  = rnorm(5,0,0.08), B = rnorm(5,0,0.08))
t2 <- data.frame(A  = rnorm(5,0,0.08), B = rnorm(5,0,0.08))
t3 <- data.frame(A  = rnorm(5,0,0.08), B = rnorm(5,0,0.08))
t4 <- data.frame(A  = rnorm(5,0,0.08), B = rnorm(5,0,0.08))
xlimits <- c(-1,1)

t1Plot <- ggplot(t1)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t1 in X")+xlim(xlimits)

t2Plot <- ggplot(t2)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t2 in X")+xlim(xlimits)

t3Plot <- ggplot(t3)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t3 in Y")+xlim(xlimits)

t4Plot <- ggplot(t4)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t4 in Y")+xlim(xlimits)

grid.arrange(t1Plot,t2Plot, t3Plot, t4Plot, ncol=2)
```
\normalsize

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
s <- function (table){ mean(table$A) - mean(table$B)}
whole <- rbind(t1,t2,t3,t4)
rawsd <- sd(c(whole$A,whole$B))
factor <- sd(c(s(t1),s(t2),s(t3),s(t4)))
numerator <-  mean(s(t1),s(t2)) - mean(s(t3),s(t4))
```
\normalsize
\pause

\footnotesize

\vspace{-2mm}

- Raw sd in data is `r round(rawsd,3)`
- The sd of means is `r round(factor,3)`
- The WEAT score is `r round(numerator/factor,3)`   
- The largest effect size reported by @Caliskan2017semanticsBiases is 1.81!

## The problem with pre-averaging on realistic set-up

### Simulation with realistic set-up (16 predicates)

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%"}
set.seed(123456)
t1 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t2 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t3 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t4 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t5 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t6 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t7 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t8 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))

xlimits <- c(-1,1)

t1Plot <- ggplot(t1)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t1 in X")+xlim(xlimits)

t2Plot <- ggplot(t2)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t2 in X")+xlim(xlimits)

t8Plot <- ggplot(t8)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t3 in Y")+xlim(xlimits)

t9Plot <- ggplot(t9)+geom_point(aes(y = 0, x = A), size = 1, alpha = 0.5)+geom_point(aes(y = 0.1, x = B), col = "skyblue", size = 1, alpha = 0.5)+theme_tufte()+xlab("similarity")+scale_y_continuous(limits = c(-.05,.15),breaks = c(0,0.1), labels = c("A","B"))+ylab("group") + geom_vline(aes(xintercept = mean(A)), size = 0.3) + geom_vline(aes(xintercept = mean(B)), size = 0.3, col = "skyblue") +ggtitle("Similarities for term t4 in Y")+xlim(xlimits)

grid.arrange(t1Plot,t2Plot, t8Plot, t9Plot, ncol=2)

whole <- rbind(t1,t2,t3,t4)
rawsd <- sd(c(whole$A,whole$B))
factor <- sd(c(s(t1),s(t2),s(t3),s(t4)))
numerator <-  mean(s(t1),s(t2)) - mean(s(t3),s(t4))
```

\footnotesize

\vspace{-2mm}

 Raw sd: `r round(rawsd,3)`, sd of means: `r round(factor,3)`, WEAT:  `r round(numerator/factor,3)`.

## The problem with pre-averaging on realistic set-up

### 10k simulations (same parameters)

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}

set.seed(123)
biasesD0 <- numeric(10000)

for(i in 1:10000){
  t1 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t2 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t3 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t4 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t5 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t6 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t7 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t8 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))
  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,0,0.08))

  factorD0 <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8),s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))
  numeratorD0 <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8)) - mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
  biasesD0[i] <- numeratorD0/factorD0
}

ggplot()+geom_histogram(aes(x=biasesD0, y = ..density..), alpha = 0.6, bins=50)+
  theme_tufte()+labs(title="10k biases for same means and sd =.08")+ xlab("bias")

```
\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
quantile <- quantile(biasesD0, c(0.025, 0.975))
absOver182 <- mean(abs(biasesD0)>=1.81)
```
\normalsize
\pause

\footnotesize

\vspace{-2mm}

- 95% of the scores are in range `r round(quantile, 3)`
- `r round(absOver182*100, 2)`% of the absolute values are above 1.81

## The problem with pre-averaging on realistic set-up

### 10k simulations with mean similarity  0.1

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
set.seed(123456)
biasesD1 <- numeric(10000)

for(i in 1:10000){
  t1 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t2 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t3 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t4 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t5 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t6 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t7 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t8 <- data.frame(A  = rnorm(8,.1,0.08), B = rnorm(8,0,0.08))
  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))
  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.1,0.08))

  factorD1 <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8),s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))
  numeratorD1 <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8)) - mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
  biasesD1[i] <- numeratorD1/factorD1
}

ggplot()+geom_histogram(aes(x=biasesD1, y = ..density..), alpha = 0.6, bins=50)+
  theme_tufte()+labs(title="10k biases for different means 0.1 and sd =.08")+ xlab("bias")

```
\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
quantiles1 <- quantile(biasesD1, c(0.025, 0.975))
absMean1 <- mean(abs(biasesD1)>=1.81)
```
\normalsize

\footnotesize

\vspace{-2mm}

- 95% of the scores are in range `r round(quantiles1, 3)`
- `r round(absMean1*100, 2)`% of the absolute values are above 1.81

## The problem with pre-averaging on realistic set-up

### 10k simulations with mean similarity  0.4

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
set.seed(123456)
biasesD4 <- numeric(10000)

for(i in 1:10000){
  t1 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t2 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t3 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t4 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t5 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t6 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t7 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t8 <- data.frame(A  = rnorm(8,.4,0.08), B = rnorm(8,0,0.08))
  t9 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t10 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t11 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t12 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t13 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t14 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t15 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))
  t16 <- data.frame(A  = rnorm(8,0,0.08), B = rnorm(8,.4,0.08))

  factorD4 <- sd(c(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8),s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16)))
  numeratorD4 <-  mean(s(t1),s(t2),s(t3),s(t4),s(t5),s(t6),s(t7),s(t8)) - mean(s(t9),s(t10),s(t11),s(t12),s(t13),s(t14),s(t15),s(t16))
  biasesD4[i] <- numeratorD4/factorD4
}

ggplot()+geom_histogram(aes(x=biasesD4, y = ..density..), alpha = 0.6, bins=50)+
  theme_tufte()+labs(title="10k biases for different means 0.4 and sd =.08")+ xlab("bias")

```
\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
quantiles4 <- quantile(biasesD4, c(0.025, 0.975))
absMean4 <- mean(abs(biasesD4)>=1.81)
```
\normalsize

\footnotesize

\vspace{-2mm}

- 95% of the scores are in range `r round(quantiles4, 3)`
- `r round(absMean4*100, 2)`% of the absolute values are above 1.81

## Advantages of the Bayesian way

- Direct impact of sample sizes
- Straightforward interpretation in terms of posterior probabilities
- Freedom to choose granularity level
- More honest risk assessment and decision making

## Bayesian model

### Choosing priors

```{r priorsVis,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, out.width = "100%", fig.caption = "Regularizing priors for the bayesian models."}
library(ggpubr)
sim <- seq(-1,1,by = 0.001)
dis <- 1-sim

p <- ggplot(data = data.frame(distance = dis), 
            mapping = aes(x = distance))+theme_tufte(base_size = 15)

none <- function(x) dnorm(x,1,.5)
cau <- function(x) dcauchy(x,0,1)
par <- function(x) dnorm(x,0,1)

noneG <- p + stat_function(fun = none)+xlim(c(-0.5,2.5))+ggtitle("Prior for mean distances")
parG <- p + stat_function(fun = par)+xlim(c(-2,2))+xlab("distance change")+ggtitle("Prior for coefficients")
cauG <- p + stat_function(fun = cau)+xlim(c(0,2))+ggtitle("Prior for standard deviation")+xlab("distance change")
ggarrange(noneG,cauG, parG, ncol = 1)
```

## Bayesian model architecture

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
library(rethinking)
options(buildtools.check = function(action) TRUE )
religionCoefs <- ulam(
  alist(
    cosineDistance ~ dnorm(mu,sigma),
    mu <- m + co[con],
    m ~ dnorm(1,.5),
    co[con] ~ dnorm(0,.5),
    sigma ~ dcauchy(0,1)
  ),
  data = religion,
  chains=2 , iter=8000 , warmup=1000, 
  log_lik = TRUE
)
```
\normalsize

##   Dataset-level coefficients

### Religion with 89\%-compatibility intervals (HPDI)

\begin{center}
\begin{figure}[!htb]\centering
   \begin{minipage}{0.55\textwidth}
  \includegraphics[width=6cm]{../../images/religionCoeffs.jpeg}
   \end{minipage} \begin{minipage}{0.4\textwidth}\footnotesize
   \begin{itemize}
   \item All HPDIs overlap with 0 
   \item Differences between classes are relatively small
   \item Coefficients for Race are similar
   \end{itemize}
   \end{minipage}
  \end{figure}

\end{center}

##  Dataset-level  coefficients

### Gender with 89\%-compatibility intervals (HPDI)

\begin{center}
\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=6cm]{../../images/genderCoeffs.jpeg}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize
   \begin{itemize}
   \item Associated and different are away from 0
   \item But they were supposed to be opposites and are very close to each other (co-occurrence?)
   \item Differences between classes are still relatively small
   \end{itemize}
   \end{minipage}
\end{figure}

\end{center}

## Bayesian model architecture

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
library(rethinking)
options(buildtools.check = function(action) TRUE )
religionCoefs <- ulam(
  alist(
    cosineDistance ~ dnorm(mu,sigma),
    mu <- m[pw] + co[con],
    m[pw] ~ dnorm(1,.5),
    co[con] ~ dnorm(0,.5),
    sigma ~ dcauchy(0,1)
  ),
  data = religion,
  chains=2 , iter=8000 , warmup=1000, 
  log_lik = TRUE
)
```
\normalsize

## Word-level coefficients

\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7.5cm]{../../images/visReligionReddit.png}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize

\vspace{-4cm}

   \begin{itemize}
   \item Most intervals overlap with control groups
   \item Often not too much difference between associated and different
   \end{itemize}
   \end{minipage}
\end{figure}

## Word-level coefficients

\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7.5cm]{../../images/visGenderReddit.png}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize

\vspace{-4cm}

   \begin{itemize}
   \item Male attributes: strong co-occurrence with female attributes
   \item Sometimes \textsf{different} is closer than \textsf{associated}
   \item Almost no overlap with control groups
   \end{itemize}
   \end{minipage}
\end{figure}

## Word-level coefficients

\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7.5cm]{../../images/visRaceReddit.png}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize

\vspace{-4cm}

   \begin{itemize}
   \item A lot of variation between races

   \item Often not much difference between associated and different
   \end{itemize}
   \end{minipage}
\end{figure}

## Thank you!

### Summary

- Bias in word embeddings
- WEAT and MAC methods
- Methodological problems
- Limitations of pre-averaging in bias detection methods
- Accounting for uncertainty with Bayesian approach

### Further work

- Including contrasts in Bayesian calculation
- Performance cross-validation in comparison to other methods (regular linear regression, KNN, \dots)
- Downstream tasks and connection with intrinsic evaluation
- Testing data from the original Implicit Association Test (IAT) 
- Applying uncertainty to WEAT and better word lists
- Looking at other similarity measures

## References
\tiny