### 2022-03-08 presentations/presentationBoston/tryOut.Rmd

---
title: '\Large Taking uncertainty seriously  \newline \normalsize A  Bayesian approach
  to  word embedding bias estimation '
author: Alicja Dobrzeniecka \& Rafal  Urbaniak \footnotesize \newline (LoPSE research
  group, University of  Gdansk)
date: "Boston, April Fools' Day"
output:
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
  slidy_presentation: default
fontsize: 10pt
classoption: x11names, dvipsnames, bibspacing, natbib, table
urlcolor: blue
bibliography: ../../references/cosineReferences1.bib
csl: ../../references/apa-6th-edition.csl
link-citations: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(ggforce)
library(ggExtra)
library(ggthemes)
library(latex2exp)
th <- theme_tufte(base_size = 20)
options(kableExtra.latex.load_packages = FALSE)
```

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine1,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Vectors}
\begin{align*}
a  & = [1,2]\\
b  &= [3,2]
\end{align*}

\end{block}
\pause

\begin{block}{Dot product}

\begin{align*}
a \cdot b & = a_1 b_1 + a_2 b_2\\
a \cdot a & = a_1^2 + a_2 ^ 2 \\
\lVert a\rVert & = \sqrt(a \cdot a)
\end{align*}

\end{block}

\end{columns}

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine2,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)
bma <- b - a

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)+
  geom_segment(aes(x = 0, y = 0, xend = bma[1], yend = bma[2]),
      arrow = arrow(length = unit(0.02, "npc")), color = "grey")+ 
  annotate(geom = "text", label  = "b - a = [2,.5]", x = 2.2, y = .9, size = 7)+
  theme_tufte(base_size = 24)

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Vectors}

\begin{align*}
a  & = [1,2]\\
b  &= [4,4]
\end{align*}

\end{block}

\begin{block}{Dot product}

\begin{align*}
a \cdot b & = a_1 b_1 + a_2 b_2\\
a \cdot a & = a_1^2 + a_2 ^ 2 \\
\lVert a\rVert & = \sqrt(a \cdot a)
\end{align*}

\end{block}

\begin{block}{Vector difference}

\begin{align*}
b - a & = [b_1- a_1, b_2 - a_2 ]
\end{align*}

\end{block}

\end{columns}

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine3,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)+
  geom_segment(aes(x = a[1], y = a[2], xend = b[1], yend = b[2]),
      arrow = arrow(length = unit(0.02, "npc")), color = "grey")+ 
  annotate(geom = "text", label  = "b - a = [2,.5] (shifted)", x = 2, y = 2.7, size = 7)+
  theme_tufte(base_size = 24)

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Vectors}

\begin{align*}
a  & = [1,2]\\
b  &= [4,4]
\end{align*}

\end{block}

\begin{block}{Dot product}

\begin{align*}
a \cdot b & = a_1 b_1 + a_2 b_2\\
a \cdot a & = a_1^2 + a_2 ^ 2 \\
\lVert a\rVert & = \sqrt(a \cdot a)
\end{align*}

\end{block}

\begin{block}{Vector difference}

\begin{align*}
b - a & = [b_1- a_1, b_2 - a_2 ]
\end{align*}

\end{block}

\end{columns}

## Cosine similarity

\begin{columns}
\column{0.45\linewidth}

```{r cosine4,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
a <- c(1,2)
b <- c(3,2.5)
#atan2(a[1] , a[2])
#atan2(b[1] , b[2])

ggplot()+
  xlim(c(0,4.5))+ylim(c(0,4.5))+
  geom_segment(aes(x = 0, y = 0, xend = a[1], yend = a[2]),  
               arrow = arrow(length = unit(0.02, "npc")))+
  annotate(geom = "text", label  = "a = [1,2]", x = 1.2, y = 2.4, size = 7)+
  geom_segment(aes(x = 0, y = 0, xend = b[1], yend = b[2]), arrow = arrow(length = unit(0.02, "npc")))+   annotate(geom = "text", label  = "b = [3,2.5]", x = 3.2, y = 2.9, size = 7)+
  theme_tufte(base_size = 24)+
  geom_segment(aes(x = a[1], y = a[2], xend = b[1], yend = b[2]),
      arrow = arrow(length = unit(0.02, "npc")), color = "grey")+ 
  annotate(geom = "text", label  = "b - a = [2,.5] (shifted)", x = 2, y = 2.7, size = 7)+
  theme_tufte(base_size = 24)+
  annotate(geom = "text", label  = expression(theta), x = .6, y = .78, size = 7)+
  geom_arc(aes(x0 = 0, y0 = 0, r = 1.3, 
               start = .463, end = .876))

```

\column{0.5\linewidth}

\footnotesize

\begin{block}{Angle}

\begin{align*}
\lVert b - a \rVert^2 &  = \lVert  b \rVert^2 + \lVert a \rVert^2 - 
  2 \lVert b \rVert   \lvert  a \rVert  \cos \theta  \\
b \cdot a  & = \lVert b\rVert \lVert  a \rVert \cos \theta \\ 
\cos \theta & = \frac{b \cdot a}{\lVert b\rVert  \lVert  a \rVert }
\end{align*}

\end{block}

\pause

\begin{block}{Orthogonality}
\begin{align*}
\cos ( 90^{\circ}) & = 0 \\ 
\frac{b \cdot a}{\lVert  b\rVert \lVert a \rVert} & = 0\\ 
 b \cdot a & = 0  
\end{align*}
\end{block}

\end{columns}