### 2022-03-08 presentations/presentationBoston/tryOut2.Rmd

---
title: '\Large Taking uncertainty seriously  \newline \normalsize A  Bayesian approach
  to  word embedding bias estimation '
author: Alicja Dobrzeniecka \& Rafal  Urbaniak \footnotesize \newline (LoPSE research
  group, University of  Gdansk)
date: "Boston, April Fools' Day"
output:
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
  slidy_presentation: default
fontsize: 10pt
classoption: x11names, dvipsnames, bibspacing, natbib, table
urlcolor: blue
bibliography: ../../references/cosineReferences1.bib
csl: ../../references/apa-6th-edition.csl
link-citations: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(ggforce)
library(ggExtra)
library(ggthemes)
library(latex2exp)
th <- theme_tufte(base_size = 20)
options(kableExtra.latex.load_packages = FALSE)
```

## Word2vec

### Question

How to sensibly represent words with numbers?

\pause

### One-hot encoding

Well, you could use 30k binary vectors with a slot for each lexical unit\dots

\pause

\dots \dots but this would be inefficient and wouldn't capture any relations between words.

## Word2vec

\begin{center}
 \includegraphics[width = 9cm]{images/perceptron1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

\footnotesize

\begin{block}{Rosenblatt's perceptron}
\begin{itemize}
\item Inputs (pixel intensities) with weights
\item Nodes with activation levels from 0-1
\item (Perhaps) 0-1 output based on a threshold
\end{itemize}
\end{block}

## Word2vec

\begin{center}
 \includegraphics[width = 9cm]{images/perceptron1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

\footnotesize

\begin{block}{Learning}
\begin{itemize}
\item Start with random weights
\item Test on a case:
\begin{itemize}
\item If right, don't change weights.
\item If wrong, change weights a bit, with focus on the ones more responsible for the judgment:
\begin{align*}
w_j & \leftarrow w_j = \overbrace{\eta}^{\text{learning rate}}(\underbrace{t}_{\text{correct output}} - \overbrace{y}^{\text{actual output}})\underbrace{x_j}_{\text{actual input}}
\end{align*}
\end{itemize}
\end{itemize}

\end{block}

## Word2vec

\begin{columns}
\column{0.45\linewidth}

\begin{center}
 \includegraphics[height = 6cm, width = 5cm]{images/perceptron2.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

\column{0.5\linewidth}

\footnotesize

\begin{itemize}
\item Each hidden unit takes a weighted sum of 324 inputs and passes on its activation level as input to outer layer units.

\item Activation levels of outer layers are interpreted as network's levels of confidence in a classification problem.

\item Learning: back-propagation (gradient descent: approximate  the direction of steepest descent in the error surface w.r.t to weights, modify accordingly).
\end{itemize}

\end{columns}

## Word2vec

### Distributional semantics

\begin{itemize}
\item "You shall know a word by the company it keeps" (John Firth, 1957)

\item "the degree of semantic similarity between two linguistic expressions $A$ and $B$ is a function of the similarity of the linguistic contexts in which $A$ and $B$ can appear." (A. Lenci, 2008)

\end{itemize}

\pause

\begin{center}
 \includegraphics[height = 5cm, width = 5cm]{images/similarity1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

## Word2vec

### Google and Mikolov

\emph{Efficient Estimation of Word Representation in Vector Space}, 2013

Let's train a neural network and use vectors of weights!

\begin{center}
 \includegraphics[height = 5cm, width = 6cm]{images/word2vec1.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

## Word2vec

### Google and Mikolov

\emph{Efficient Estimation of Word Representation in Vector Space}, 2013

Let's train a neural network and use vectors of weights!

\begin{center}
 \includegraphics[height = 3cm, width = 7cm]{images/word2vec2.png}
\end{center}

\vspace{-3mm}
\tiny \hfill \color{gray}Illustration: M. Mitchell \color{black}

## Word2vec

### Nearest words

\begin{itemize}
\item \textbf{philosophy}: philosophies, credo, ethos, principles, ethic, tenets, mantra, ideology, mindset, worldview

\item \textbf{sandwich}: sandwiches, burger, chicken sandwich, cheeseburger, burrito, burgers, pizza, turkey sandwich, hamburger, burritos
\end{itemize}

\pause

### Some similarities from \emph{philosophy}

Logic (.47), Nietzsche (.32), Hegel (.32), analytic (.13), burger (.08), continental (.04), Russell (.04)