### 2022-04-27 paperDraft/metaTechnicalDetails.Rmd
Commits old fiels

---
title: ""
author: ""
output:
  pdf_document:
    number_sections: true
    df_print: kable 
    keep_tex: true
    includes:
      in_header:
        - Rafal_latex6.sty
classoption: dvipsnames,enabledeprecatedfontcommands
fontsize: 12pt
documentclass: scrartcl
urlcolor: blue
csl: transactions-on-computational-logic.csl
---

# Recreating Code for MAC Manzini

Word embeddings sources:
\newline

- the same as the authors - Reddit-L2 corpus (reddit.l2.tar.gz) from http://cl.haifa.ac.il/projects/L2/ 
\newline

- our idea - GoogleNews corpus (GoogleNews-vectors-negative300.bin.gz.) from  https://code.google.com/archive/p/word2vec/

Getting cosine similarity and cosine distance:
\newline

Description of the precedure:

TODO: Run code with fresh downloaded corpora

# Adding control groups

## Human words assumptions and procedure of creation

TODO

## Neutral words assumptions and procedure of creation

TODO